---
title: "Genome and Transcriptome analyses of _Marteilia sydneyi_"
author: "Niki Nenadic and Ido Bar"
date: "04 July 2023"
always_allow_html: yes
output: 
    bookdown::html_document2:
      df_print: paged
      theme: united
      highlight: pygments
      css: "style/style.css"
      toc: true
      toc_float: true
      toc_depth: 3
  #    highlight: pygments
      number_sections: false
      code_folding: hide
#      keep_md: true
#bibliography: style/comparative_genomics.bib
bibliography: ["style/QX_references.bib","style/comparative_genomics.bib", 'style/oyster_genomes.bib', 'style/pseudo_counts.bib', 'style/stLFR_assembers.bib', 'style/transcriptome_assembly_evaluation.bib', 'style/Metagenomes_assembly_papers.bib', 'style/NGS_QC.bib', 'style/RNAseq_tools.bib'] 
csl: style/springer-basic-improved-author-date-with-italic-et-al-period.csl
---

```{r setup, include=FALSE}
devtools::source_gist("7f63547158ecdbacf31b54a58af0d1cc", filename = "util.R")
knitr::opts_chunk$set(list(echo = TRUE, eval=FALSE, message=FALSE))
fav_packages <- c('tidyverse', 'rentrez', 'taxonomizr')
pak::pak(fav_packages)

pacman::p_load(char = basename(fav_packages))
```

# Experimental Design

## Aims

-   Genomic and transcriptomic characterisation of _Marteilia sydneyi_
-   Identify genes that play a role in the defence response of Sydney Rock Oysters (_Saccostrea glomerata_) during infection of _Marteilia sydneyi_ (QX disease)  

### Objectives

1. Assemble the _Marteilia sydneyi_ transcriptome
2. Identify differential gene expression in infected and non-infected Sydney Rock Oysters (_Saccostrea glomerata_)
3. Assemble the _Marteilia sydneyi_ genome

# Transcriptome assembly of _M. sydneyi_ and _Saccostrea glomerata_

## Analysis Pipeline

### General overview:

1.  Data pre-processing:
    a.  Quality check
    b.  Adaptor trimming
    c.  Post-trim quality check
    d.  Error correction (?)
    e.  Combine reads from control SRO and QX-infected SRO
2.  Mapping reads to host reference genome
3.  _De Novo_ transcriptome assembly of the reads  
4.  Homology-based taxonomic classification of transcripts  
5.  Separation of host, _M. sydneyi_ and off-target (various contaminations) transcripts
6.  Annotation of host pathogen transcripts
7.  Host transcriptome differential expression
8.  Summary statistics and visualisation

### Methods
#### RNA extraction
RNA was extracted from the digestive gland of 3 healthy Sydney Rock Oysters (SRO, _Saccostrea glomerata_) and 3 SRO infected with QX disease (as confirmed by PCR assay on DNA extracted from each tissue) using the Trizol method. The RNA was sent for Illumina paired-end sequencing, generating 150bp reads (Macrogen, Korea). 

The reads were downloaded to the Griffith University High Performance Computing (HPC) cluster (Gowonda) for bioinformatics processing and analyses, following the guidelines by Harvard Informatics (see [link](https://github.com/harvardinformatics/TranscriptomeAssemblyTools)) . 
These included error-correction with Rcorrect v1.5.0 [@songRcorrectorEfficientAccurate2015a] of the reads from each set (healthy SRO and QX-infected SRO) and removal of "unfixable" reads.

```{bash prep-rnaseq-reads}
# setup environment
CONDA_NAME="genomics"
mamba install -n $CONDA_NAME rcorrector agat # can be in the base environment
cd $HOME/etc/tools/Trinity # ~/aDNA/tools 
# install additional utilities
git clone https://github.com/harvardinformatics/TranscriptomeAssemblyTools.git
WORKDIR=/scratch/aDNA/niki_qx
mkdir -p $WORKDIR
cd $WORKDIR
# Prepare a general array PBS script
echo '#!/bin/bash 
#PBS -V

cd $WORKDIR
>&2 echo Current working directory: $PWD
source ~/.bashrc
conda activate $CONDA_NAME 
set -Eeo pipefail'"
gawk -v ARRAY_IND=\$PBS_ARRAY_INDEX 'NR==ARRAY_IND' \$CMDS_FILE | bash" > ${WORKDIR}/array.pbspro
# download reads
rclone copy -P GU_owncloud:Shared/Oyster_QX_Disease/QX_RNA-Seq ./QX_RNA-Seq
cd QX_RNA-Seq/data && mkdir corrected_reads

# prepare commands to run Rcorrector for each sample in parallel
# ls -1 NN01-5*_1.fastq.gz | parallel -k --dry-run --rpl "{infile2} s:_1:_2:;" "run_rcorrector.pl -1 {} -2 {infile2} -verbose -t \$NCPUS -od corrected_reads"  > rcorrector.cmds
# submit the jobs
# qsub -J1-$(cat rcorrector.cmds | wc -l) -l select=1:ncpus=8:mem=16GB,walltime=10:00:00 -N rcorr_batch -v "CMDS_FILE=rcorrector.cmds","WORKDIR=$PWD","CONDA_NAME=$CONDA_NAME" ${WORKDIR}/array.pbspro
# restart failed jobs at stage 3
# ls -1 NN01-5*_1.fastq.gz | parallel -k --dry-run --rpl "{infile2} s:_1:_2:;" "run_rcorrector.pl -1 {} -2 {infile2} -t \$NCPUS -od corrected_reads -stage 3" > rcorrector_s3.cmds
# submit the jobs
# qsub -J1-$(cat rcorrector_s3.cmds | wc -l) -l select=1:ncpus=8:mem=24GB,walltime=10:00:00 -N rcorr_batch_s3 -v "CMDS_FILE=rcorrector_s3.cmds","WORKDIR=$PWD","CONDA_NAME=$CONDA_NAME" ${WORKDIR}/array.pbspro
# didn't work for NN-567 for some reason
# restart job for NN-567 from stage 1
# echo "source ~/.bashrc; conda activate $CONDA_NAME; cd $PWD; set -Eeo pipefail; run_rcorrector.pl -1 NN01-567_1.fastq.gz -2 NN01-567_2.fastq.gz -verbose -t \$NCPUS -od corrected_reads" | qsub -l select=1:ncpus=12:mem=48gb,walltime=20:00:00 -N rcorrect_567 -q bigmem

# Run Rcorrect on Control samples
mkdir corrected_reads
R1_FILES_CON=$(seq 1 3 | parallel -k echo "NN01-56{}_1.fastq.gz" | paste -sd,)
R2_FILES_CON=$(seq 1 3 | parallel -k echo "NN01-56{}_2.fastq.gz" | paste -sd,)
# submit job
echo "source ~/.bashrc; conda activate $CONDA_NAME; cd $PWD; set -Eeo pipefail; run_rcorrector.pl -1 $R1_FILES_CON -2 $R2_FILES_CON -t \$NCPUS -od corrected_reads" | qsub -l select=1:ncpus=12:mem=48gb,walltime=50:00:00 -N rcorr_sro_control -q bigmem  

# combine and correct QX-infected samples
R1_FILES_QX=$(printf "67\n79\n80\n" | parallel -k echo "NN01-5{}_1.fastq.gz" | paste -sd,)
R2_FILES_QX=$(printf "67\n79\n80\n" | parallel -k echo "NN01-5{}_2.fastq.gz" | paste -sd,)
# submit job
echo "source ~/.bashrc; conda activate $CONDA_NAME; cd $PWD; set -Eeo pipefail; run_rcorrector.pl -1 $R1_FILES_QX -2 $R2_FILES_QX -t \$NCPUS -od corrected_reads" | qsub -l select=1:ncpus=12:mem=48gb,walltime=50:00:00 -N rcorr_sro_qx -q bigmem 

# combine reads from individually-corrected samples
cd corrected_reads
echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; seq 1 3 | parallel -k -j4 cat NN01-56{}_1.cor.fq.gz > SRO_control_R1.cor.fastq.gz ; seq 1 3 | parallel -k -j4 cat NN01-56{}_2.cor.fq.gz > SRO_control_R2.cor.fastq.gz ; "'printf "67\n79\n80\n" | parallel -k -j4 cat NN01-5{}_1.cor.fq.gz > SRO_QX_R1.cor.fastq.gz ; printf "67\n79\n80\n" | parallel -k -j4 cat NN01-5{}_2.cor.fq.gz > SRO_QX_R2.cor.fastq.gz' | qsub -l select=1:ncpus=8:mem=24gb,walltime=3:00:00 -N combine_corr_reads 

# remove unfixable reads
# SRO control
echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; python  ~/etc/tools/Trinity/TranscriptomeAssemblyTools/utilities/FilterUncorrectabledPEfastq.py -1 SRO_control_R1.cor.fastq.gz -2 SRO_control_R2.cor.fastq.gz -s SRO_control " | qsub -l select=1:ncpus=8:mem=24gb,walltime=5:00:00 -N rm_unfixable_control
# SRO-QX
echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; python  ~/etc/tools/Trinity/TranscriptomeAssemblyTools/utilities/FilterUncorrectabledPEfastq.py -1 SRO_QX_R1.cor.fastq.gz -2 SRO_QX_R2.cor.fastq.gz -s SRO_QX" | qsub -l select=1:ncpus=8:mem=24gb,walltime=5:00:00 -N rm_unfixable_qx

# compress files with genozip so we could upload them
ls -1 *.fastq.gz | parallel --dry-run genozip -@ \$NCPUS -f {} > genozip_files.cmds
# submit the jobs
qsub -J1-$(cat genozip_files.cmds | wc -l) -l select=1:ncpus=8:mem=16GB,walltime=4:00:00 -N genozip_fastq -v "CMDS_FILE=genozip_files.cmds","WORKDIR=$PWD","CONDA_NAME=base" ${WORKDIR}/array.pbspro

# upload files to Research Space
rclone copy -P $WORKDIR/QX_RNA-Seq/data GU_owncloud:Shared/Oyster_QX_Disease/QX_RNA-Seq/data
```


```{bash rcorrect-bunya}
WORKDIR=/scratch/project/adna/
mkdir -p $WORKDIR/QX_RNA-Seq/data/combined_reads
cd $WORKDIR/QX_RNA-Seq/data
# prepare slurm script
echo '#!/bin/bash --login
#SBATCH --job-name=QX-combine-reads
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log
#SBATCH --cpus-per-task=8
#SBATCH --mem=16G
#SBATCH --time=4:00:00
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
set -Eeo pipefail
cd $SLURM_SUBMIT_DIR
srun seq 1 3 | parallel -k cat NN01-56{}_1.fastq.gz > combined_reads/SRO_control_R1.fastq.gz; \
seq 1 3 | parallel -k cat NN01-56{}_2.fastq.gz > combined_reads/SRO_control_R2.fastq.gz; \
printf "67\n79\n80\n" | parallel -k cat NN01-5{}_1.fastq.gz > combined_reads/SRO_QX_R1.fastq.gz; \
printf "67\n79\n80\n" | parallel -k cat NN01-5{}_2.fastq.gz > combined_reads/SRO_QX_R2.fastq.gz' > combine_reads.slurm
# submit the job 
sbatch combine_reads.slurm

echo '#!/bin/bash --login
#SBATCH --job-name=QX-combine-rcorrect
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log
#SBATCH --cpus-per-task=8
#SBATCH --mem=16G
#SBATCH --time=4:00:00
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
set -Eeo pipefail
cd $SLURM_SUBMIT_DIR
srun seq 1 3 | parallel -k cat NN01-56{}_1.cor.fq.gz > SRO_control_R1.cor.fastq.gz; \
seq 1 3 | parallel -k cat NN01-56{}_2.cor.fq.gz > SRO_control_R2.cor.fastq.gz; \
printf "67\n79\n80\n" | parallel -k cat NN01-5{}_1.cor.fq.gz > SRO_QX_R1.cor.fastq.gz; \
printf "67\n79\n80\n" | parallel -k cat NN01-5{}_2.cor.fq.gz > SRO_QX_R2.cor.fastq.gz' > combine_cor_reads.slurm
# submit the job 
sbatch combine_cor_reads.slurm

# remove unfixable reads
# SRO control
echo '#!/bin/bash --login
#SBATCH --job-name=rm_unfixable_control
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log
#SBATCH --cpus-per-task=8
#SBATCH --mem=24G
#SBATCH --time=6:00:00
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
set -Eeo pipefail
cd $SLURM_SUBMIT_DIR
srun python  ~/etc/tools/TranscriptomeAssemblyTools/utilities/FilterUncorrectabledPEfastq.py -1 SRO_control_R1.cor.fastq.gz -2 SRO_control_R2.cor.fastq.gz -s SRO_control ' > rm_unfixable_control.slurm
# submit the job 
sbatch rm_unfixable_control.slurm

# SRO-QX
echo '#!/bin/bash --login
#SBATCH --job-name=rm_unfixable_qx
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log
#SBATCH --cpus-per-task=8
#SBATCH --mem=24G
#SBATCH --time=6:00:00
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
set -Eeo pipefail
cd $SLURM_SUBMIT_DIR
srun python  ~/etc/tools/TranscriptomeAssemblyTools/utilities/FilterUncorrectabledPEfastq.py -1 SRO_QX_R1.cor.fastq.gz -2 SRO_QX_R2.cor.fastq.gz -s SRO_QX' > rm_unfixable_qx.slurm
# submit the job 
sbatch rm_unfixable_qx.slurm

```

#### Transcriptome assembly and Annotation with TransPi
A _De-Novo_ transcriptome was assembled from the combined error-corrected reads using [TransPi](https://palmuc.github.io/TransPi) [@rivera-vicensTransPiComprehensiveTRanscriptome2022]. In brief, TransPi performs a complete transcriptome assembly, basic annotation and assessment pipeline using well-established methods:  

 * Trimming of low-quality bases and removal of sequencing adaptors with Fastp [@chenFastpUltrafastAllone2018]  
 * Ribosomal RNA (rRNA) removal matching against the [SILVA rRNA database](https://www.arb-silva.de/) [@quastSILVARibosomalRNA2012]  
 * Transcriptome assembly using a range of tools, including Trinity, SPAdes, Trans-ABySS and SOAPdenovo-Trans, with a range of k-mers (see a comprehensive comparison of these tools at @holzerNovoTranscriptomeAssembly2019 and of k-mer selection at @haznedarogluOptimizationNovoTranscriptome2012)   
 * Collapsing redundant transcripts with [EvidentialGene](http://arthropods.eugenes.org/EvidentialGene/)  
 * Annotation of the assemblies against UniProt databases  
 * Building a [Trinotate](https://github.com/Trinotate/Trinotate/wiki) database for each assembly and its annotation  
 
TransPi needs to be installed and configured (download annotation databases and set their paths in a `nextflow.config` file). Unfortunately, a lot of the configuration for TransPi to work on GU Gowonda needs to be done manually, including downloading the Singularity containers and setting up the resources required for each process - see the current configuration and the following reported issues:  

1. Missing BUSCO container in the configuration - [Issue #53](https://github.com/PalMuc/TransPi/issues/53#issuecomment-1268051333)  
2. The precheck script fails to download the UniProt database due to change in the API - see [Issue #52](https://github.com/PalMuc/TransPi/issues/52)  
3. Use the single TransPiContainer profile to fix issues with some of the tools (see the [documentation](https://palmuc.github.io/TransPi/#_containers))  
4. Trinity and Velvet-Oases jobs are failing though they seem to complete without a problem (no output files are copied across from the compute nodes). It can be one of these possibilities:  
    a. Maybe I simply ran out of space (in `/scratch/project/adna`), I've requested additional storage on the `adna` folder on Bunya (now I have 10TB there)  
    b. It may be related to the use of `$TMPDIR` on the compute nodes (change the flag `scracth=true` into `scratch='/scratch/project/adna/tmp'` in `~/.nextflow/bunya.config`)  
    c. `$SINGULARITY_TMPDIR` that is running out of space or not properly configured, see this [thread on GitHub](https://github.com/nextflow-io/nextflow/issues/2786#issuecomment-1173975771)  
    d. Another option is that Bunya is using Apptainer and not Singularity, so maybe I need to configure a profile to use Apptainer instead (v22.11.1).  
    e. Found the problem!!! It appears that the way TransPi is writing the version numbers throws an error (when `$v` is defined multiple times). I fixed it by using `printf` statement instead of the multiple `echo` ones. 
    f. Some processes (such as `summary_custom_uniprot`) fail due to `set -o pipefail` directive (complex pipelines, including `grep`, etc.). It can be fixed by adding `set +o pipefail` to the top of the script.  
5. `Rnammer` fails - it needs to be setup to use `hmmsearch2` (which can be downloaded from [this link](http://eddylab.org/software/hmmer/hmmer-2.3.2.tar.gz) and remove the `--cpu 1` flag from `core-rnammer`, see details [here](https://groups.google.com/g/trinityrnaseq-users/c/WZjkGSMUT3I))  
6. `rnaQUAST` fails -- it requires an additional tool (GeneMark S-T) to be installed separately and put in the container's `$PATH` (add `export PATH=\$PATH:\$HOME/etc/tools/Annotation/GeneMarkST/` at the top of the script of the rnaquast process in the main nextflow). See details [here](https://github.com/ablab/rnaquast/issues/5#issuecomment-823996456).  
7. `signalp` fails -- need to edit the executable to allow it to find where it is being run from and increase the sequence limit (see [here](https://www.seqanswers.com/forum/bioinformatics/bioinformatics-aa/29132-how-to-increase-sequence-limit-in-signalp#post236326)) with the following command:  
`sed -ri.bak 's|SIGNALP} = '\''.+|SIGNALP} = "\$FindBin::RealBin"|; s/BEGIN/use FindBin\;
\nBEGIN/; s/MAX_ALLOWED_ENTRIES=.+/MAX_ALLOWED_ENTRIES=2000000;/' signalp`
8. Process `summary_custom_uniprot` fails because of weird table merging and file redundancies done in bash. I edited `Tranbspi/bin/custom_uniprot_hits.R` to perform these tasks and export the `csv` file with the results.  

```{bash setup-transpi}
# install TransPi
cd /home/ibar/adna/tools #Bunya
git clone https://github.com/idobar/TransPi.git # my version
bash ./TransPi/precheck_TransPi.sh $PWD/TransPi
sed -i "s/'singularity' || workflow.containerEngine == 'docker'/'singularity' || workflow.containerEngine == 'docker' || workflow.containerEngine == 'apptainer'/g" $PWD/TransPi/TransPi.nf # add support for Apptainer (it will also require NXF_VER=22.11.1-edge)  
# download the pipeline containers
grep "singularity_pull_docker_container" $PWD/TransPi/TransPi.nf | cut -f 2 -d '"' | sort | uniq | tr -d \' | parallel --dry-run --rpl  "{outfile} s=https://==; s=[:/]+=-=g;" aria2c -c {} -o $NXF_SINGULARITY_CACHEDIR/{outfile}.img 

# Install GeneMark
mkdir -p ~/etc/tools/Annotation/GeneMarkST
cd  !$
aria2c -x5 http://topaz.gatech.edu/GeneMark/tmp/GMtool_ZozF5/gmst_linux_64.tar.gz
tar xzf gmst_linux_64.tar.gz
ln -s $PWD/gmst.pl ~/bin/ # add it to the PATH
# need to apply for a licence at http://topaz.gatech.edu/GeneMark/license_download.cgi
aria2c http://topaz.gatech.edu/GeneMark/tmp/GMtool_ZozF5/gm_key_64.gz

# fix Rnammer
sed -i.bak 's/--cpu 1 //g' /home/ibar/scratch/tools/rnammer-1.2/core-rnammer
# fix signalp
sed -ri.bak 's|SIGNALP} = '\''.+|SIGNALP} = "\$FindBin::RealBin"|; s/BEGIN/use FindBin\;
\nBEGIN/; s/MAX_ALLOWED_ENTRIES=.+/MAX_ALLOWED_ENTRIES=2000000;/' /home/ibar/scratch/tools/signalp-4.1/signalp

cd ~/bin
aria2c -x5 https://github.com/nextflow-io/nextflow/releases/download/v21.10.4/nextflow-21.10.4-all # this is the original version in TransPi installer - it does not support Apptainer!
chmod +x nextflow-21.10.4-all
aria2c -x5 https://github.com/nextflow-io/nextflow/releases/download/v22.11.1-edge/nextflow-22.11.1-edge-all
chmod +x nextflow-22.11.1-edge-all # this is the only version that supports DSL1, Apptainer and has an update Tower plugin!
```

##### Test TransPi on subset of reads from one sample

```{bash transpi-test}
# install bbtools
mamba install -n gwas bbtools
mkdir -p ~/adna/sandbox/TransPi_test
cd !$

echo '#!/bin/bash --login
#SBATCH --job-name=downsample_reads
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log
#SBATCH --cpus-per-task=8
#SBATCH --mem=24G
#SBATCH --time=6:00:00
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
conda activate gwas
set -Eeo pipefail
cd $SLURM_SUBMIT_DIR
srun reformat.sh  in=~/adna/Oyster_QX_Disease/QX_RNA-Seq/data/NN01-561_#.fastq.gz out=NN01-561_#.downsampled.fastq.gz samplerate=0.2' > downsample_reads.slurm
# submit the job 
sbatch downsample_reads.slurm
# conda activate nf21 # nextflow version 21.10.6
nextflow-22.11.1-edge-all run /home/ibar/adna/tools/TransPi/TransPi.nf -with-tower --all -profile apptainer,TransPiContainer,bunya -c /home/ibar/.nextflow/bunya.config --k 25,35 --maxReadLen 150 --pipeInstall /scratch/project/adna/tools/TransPi --allBuscos --buscoDist --reads './NN01-561_[1,2].downsampled.fastq.gz'

```
 
##### Run TransPi on the combined and corrected reads from each experimental group 

```{bash transpi-assemble}
# WORKDIR=/scratch/s2978925/aDNA/niki_qx #Gowonda
WORKDIR=/home/ibar/adna/Oyster_QX_Disease # Bunya
# mamba create -n nf-dsl1 nextflow=22.10.6 # need an older version (<22) to avoid Apptainer warnings
# export NXF_DEFAULT_DSL=1 # only needed if using nextflow version>22.03.0-edge and <22.12.0-edge
# NXF_VER=22.11.1-edge

# export TOWER_ACCESS_TOKEN=eyJ0aWQiOiA0NDYxfS5hNzkzY2M3NTI1YjZkZGFlMjVlZjJiZmIxNTU1MDAyMGE4YjNjZWQ0 # only needed if it is not already set in ~/.bashrc
# export NXF_VER=22.04.5
# prepare the folder and input files
mkdir -p $WORKDIR/SRO_TransPi_assembly/input_files && cd $WORKDIR/SRO_TransPi_assembly
ln -s $WORKDIR//QX_RNA-Seq/data/corrected_reads/unfixrm_SRO_*.gz input_files/
ln -s $WORKDIR//QX_RNA-Seq/data/combined_reads/*.fastq.gz input_files/
# start TransPi (Bunya) [optional tools and databases configured in nextflow.config]
nextflow-22.11.1-edge-all run /home/ibar/adna/tools/TransPi/TransPi.nf -with-tower --all -profile apptainer,TransPiContainer,bunya -c /home/ibar/.nextflow/bunya.config --k 25,35,41,53,65,71,83 --maxReadLen 150 --pipeInstall /scratch/project/adna/tools/TransPi --allBuscos --buscoDist --reads './input_files/*_R[1,2]*.fastq.gz' -resume

# start the run (Gowonda)
# nextflow run $WORKDIR/TransPi/TransPi.nf -with-tower --all --reads './input_files/*_R[1,2]*.fastq.gz'  -profile singularity,TransPiContainer,gowonda -c $HOME/.nextflow/GU.config --rRNAfilter --rRNAdb $WORKDIR/TransPi/DBs/rrna_db/SILVA_138.1_Small_and_Large_SURef_NR99_tax_silva_trunc.fasta --withSignalP --signalp $HOME/etc/tools/Annotation/signalp-4.1/signalp --withTMHMM --tmhmm $HOME/etc/tools/Annotation/tmhmm-2.0c/bin/tmhmm --withRnammer --rnam $HOME/etc/tools/Annotation/rnammer-1.2/rnammer --k 25,35,41,53,65,71,83 --maxReadLen 150 --pipeInstall $WORKDIR/TransPi --allBuscos --buscoDist -resume

```

##### Run EvidentialGene step of TransPi to collapse redundant transcripts

```{bash transpi-evigene}
JOBNAME="onlyEvigene"
NCORES=4
MEM=24
WALLTIME=20:00:00
mkdir -p $WORKDIR/SRO_TransPi_assembly/Evigene_analysis/
cd !$
ls -1 $WORKDIR/SRO_TransPi_assembly/results/assemblies/*SOAP.fa | parallel --dry-run --rpl "{ass} s:.+/(.+).SOAP.fa:\1:" --rpl "{ass_dir} s:.+/(.+).SOAP.fa:$WORKDIR/SRO_TransPi_assembly/Evigene_analysis/\1:" "source ~/.bash_profile; mkdir -p {ass_dir}/onlyEvi && cd {ass_dir}; ln -s $WORKDIR/SRO_TransPi_assembly/results/assemblies/{ass}*[!k0-9].fa {ass_dir}/onlyEvi/; nextflow-22.11.1-edge-all run /home/ibar/adna/tools/TransPi/TransPi.nf -name {ass}_evigene -with-tower --onlyEvi --outdir {ass}_evigene_results -profile apptainer,TransPiContainer,bunya -c /home/ibar/.nextflow/bunya.config" > $JOBNAME.cmds

# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,PATH=$PATH,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=base ${WORKDIR}/array.slurm

# resume (if failed)
ls -1 $WORKDIR/SRO_TransPi_assembly/results/assemblies/*SOAP.fa | parallel --dry-run --rpl "{ass} s:.+/(.+).SOAP.fa:\1:" "source ~/.bash_profile; cd $WORKDIR/SRO_TransPi_assembly/Evigene_analysis/{ass};  nextflow-22.11.1-edge-all run /home/ibar/adna/tools/TransPi/TransPi.nf -name {ass}_evigene -with-tower --onlyEvi --outdir {ass}_evigene_results -profile apptainer,TransPiContainer,bunya -c /home/ibar/.nextflow/bunya.config -resume" > $JOBNAME.resume.cmds
# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.resume.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.resume.cmds,CONDA_NAME=base ${WORKDIR}/array.slurm

# Create 

```

##### Annotate chosen assemblies

The assemblies were annotated with TransPi, using homology searches (Blast and Diamond) against the Uniprot_Sprot protein database, as well as using HMM search against the PFam database of protein families. In addition, signal peptides, ribosomal genes and transmembrane topology were predicted with SignalP, Rnammer and TMHMM, respectively.

```{bash transpi-annot}
WORKDIR=/home/ibar/adna/Oyster_QX_Disease # Bunya
mkdir -p $WORKDIR/SRO_TransPi_assembly/Annotation/onlyAnn
ln -s $WORKDIR/SRO_TransPi_assembly/results/assemblies/*Trinity.fa $WORKDIR/SRO_TransPi_assembly/Annotation/onlyAnn/
cd $WORKDIR/SRO_TransPi_assembly/Annotation

nextflow-22.11.1-edge-all run /scratch/project/adna/tools/TransPi/TransPi.nf -with-tower --onlyAnn --outdir Annotation_results -profile apptainer,TransPiContainer,bunya -c /home/ibar/.nextflow/bunya.config --withSignalP --withTMHMM --withRnammer

# create gff and gtf files from selected assemblies
JOBNAME="gtf2gff"
NCORES=4
MEM=24
WALLTIME=20:00:00
cd $WORKDIR/SRO_TransPi_assembly/reslts/assemblies

find $WORKDIR/SRO_TransPi_assembly/results/assemblies -name "*Trinity.fa" | parallel --dry-run "perl ~/etc/tools/trinityrnaseq-v2.11.0/util/misc/cdna_fasta_file_to_transcript_gtf.pl {} > {.}.gtf; agat_convert_sp_gxf2gxf.pl -g {.}.gtf -o {.}.gff" > $JOBNAME.cmds

# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,PATH=$PATH,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=genomics ${WORKDIR}/array.slurm 

rclone copy -P --dry-run --include "*Trinity*" $WORKDIR/SRO_TransPi_assembly/results/assemblies GURS_shared:Oyster_QX_Disease/QX_RNA-Seq/SRO_TransPi_assembly/results/assemblies
```

###### Annotate chosen assemblies with BLAST

In addition to the annotation performed with TransPi, the transcriptomes were annotated against the non-redundant nucleotide and protein databases of the NCBI (`nt` and `nr`, respectively) to achieve more accurate species-specific annotations.

```{bash trans-blast}
CONDA_NAME=genomics
mamba install -n $CONDA_NAME gsutil awscli blast
# download NCBI databases
mkdir -p ~/adna/tools/ncbi_db
cd ~/adna/tools/ncbi_db
WORKDIR=/scratch/project/adna/Oyster_QX_Disease
JOBNAME="update_blast_db"
NCORES=8
MEM=16
WALLTIME=20:00:00
parallel --dry-run "update_blastdb.pl --source ncbi --decompress --force --blastdb_version 5 --num_threads 0 {}" ::: nt nr taxdb > $JOBNAME.cmds
# or do it manually 
# curl -s -l ftp://ftp.ncbi.nlm.nih.gov/blast/db/v5/ | grep -E "^nr." | parallel --dry-run aria2c -x5 -c --auto-file-renaming=false https://ftp.ncbi.nlm.nih.gov/blast/db/v5/{} > download_blast_nr.cmds


# Prepare a general array Slurm script
echo '#!/bin/bash --login
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%A.%a.log'"
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

set -Eeo pipefail
source ~/.bashrc
conda activate \$CONDA_NAME
cd \$SLURM_SUBMIT_DIR
srun gawk -v ARRAY_IND=\$SLURM_ARRAY_TASK_ID 'NR==ARRAY_IND' \$CMDS_FILE | bash" > ${WORKDIR}/array.slurm


# send to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME   ${WORKDIR}/array.slurm 

# Blastn - this takes too long, see using blast-nf below
WORKDIR=/home/ibar/adna/Oyster_QX_Disease
cd $WORKDIR/SRO_TransPi_assembly/Annotation
mkdir -p $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/blastn_tax
JOBNAME="SRO_trans_blastn"
NCORES=24
MEM=96
WALLTIME=150:00:00
ls -1 $WORKDIR/SRO_TransPi_assembly/results/assemblies/*Trinity.fa | parallel --dry-run "blastn -task blastn -db /scratch/project/adna/tools/ncbi_db/nt -query {} -outfmt '6 std stitle staxids ssciname scomname' -evalue 1e-10 -max_target_seqs 20 -num_threads \$SLURM_CPUS_PER_TASK -out Annotation_results/blastn_tax/{/.}.nt_tax.outfmt6" > $JOBNAME.cmds

# submit the commands in the login node
bash $JOBNAME.cmds
# or send to the cluster as a job array
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME --dependency=afterok:8070286  ${WORKDIR}/array.slurm 

# annotate proteins - this takes too long, see using blast-nf below
JOBNAME="SRO_trans_blastp"
ls -1 $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/transdecoder/*.transdecoder.pep | parallel --dry-run "blastp -task blastp -db /scratch/project/adna/tools/ncbi_db/nr -query {} -outfmt '6 std stitle staxids ssciname scomname' -evalue 1e-10 -max_target_seqs 20 -num_threads \$SLURM_CPUS_PER_TASK -out Annotation_results/blastp_tax/{/.}.nr_tax.outfmt6" > $JOBNAME.cmds

# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME --dependency=afterok:8070286  ${WORKDIR}/array.slurm

```

Regular BLAST is just too slow (mainly on the protein level), so we used a Nextflow implementation of BLAST that breaks down the input transcriptome into "chunks" and processes them in parallel on the HPC cluster.

```{bash blast-nf}
WORKDIR=/home/ibar/adna/Oyster_QX_Disease
cd $WORKDIR/SRO_TransPi_assembly/Annotation
JOBNAME=blastn_tax_nf
NCORES=4
MEM=24
WALLTIME=150:00:00
mkdir -p $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/$JOBNAME
ls -1 $WORKDIR/SRO_TransPi_assembly/results/assemblies/*Trinity.fa | parallel --dry-run --rpl "{sample} s:.+/(.+)\.Trinity.+:\1:" "~/bin/nextflow-22.11.1-edge-all run ~/etc/tools/blast-nf/main.nf --app blastn --dbDir /scratch/project/adna/tools/ncbi_db --dbName nt --query {} --outfmt \"'6 std stitle staxids ssciname scomname'\" --options '-evalue 1e-10 -max_target_seqs 20' --chunkSize 2500 --outdir  $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/$JOBNAME/{sample} --outfileName {/.}.nt_tax.outfmt6  -c ~/.nextflow/bunya.config -profile bunya,apptainer -with-tower" > $JOBNAME.cmds

# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=base ${WORKDIR}/array.slurm

JOBNAME=blastp_tax_nf
mkdir -p $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/$JOBNAME
ls -1 $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/transdecoder/*.transdecoder.pep | parallel --dry-run --rpl "{sample} s:.+/(.+)\.Trinity.+:\1:" "~/bin/nextflow-22.11.1-edge-all run ~/etc/tools/blast-nf/main.nf --app blastp --dbDir /scratch/project/adna/tools/ncbi_db --dbName nr --query {} --outfmt \"'6 std stitle staxids ssciname scomname'\" --options '-evalue 1e-10 -max_target_seqs 20' --chunkSize 500 --outdir  $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/$JOBNAME/{sample} --outfileName {/}.nr_tax.outfmt6  -c ~/.nextflow/bunya.config -profile bunya,apptainer -with-tower" > $JOBNAME.cmds

# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=base ${WORKDIR}/array.slurm

```

###### Annotate predicted proteins with DIAMOND

Even when using a split-combine strategy using `blast-nf`, BLASTp is just too slow (it takes around 50 hours to process a chunk of 500 proteins). Therefore we also used DIAMOND, which can be x1,000 quicker than BLAST [@buchfinkFastSensitiveProtein2015a].


```{bash prot-diamond}
WORKDIR=/home/ibar/adna/Oyster_QX_Disease
cd /scratch/project/adna/tools
JOBNAME=prep_diamond_db
NCORES=8
MEM=48
WALLTIME=10:00:00

DB_PATH="/scratch/project/adna/tools/ncbi_db"
DB="nr"
# prepare script
echo '#!/bin/bash --login
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%A.%a.log'"
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

set -Eeo pipefail
source ~/.bashrc
conda activate \$CONDA_NAME
cd \$SLURM_SUBMIT_DIR

blastdbcmd -entry all -db ${DB_PATH}/${DB} -out ${DB_PATH}/${DB}.fasta 
printf \"Finished extracting sequences from ${DB} database as '${DB}.fasta'\n\"
aria2c -x5 -c --auto-file-renaming=false -d ${DB_PATH}  https://ftp.ncbi.nlm.nih.gov/pub/taxonomy/accession2taxid/prot.accession2taxid.FULL.gz
aria2c -x5 -c --auto-file-renaming=false -d ${DB_PATH}  https://ftp.ncbi.nlm.nih.gov/pub/taxonomy/taxdmp.zip
cd ${DB_PATH} && unzip taxdmp.zip
gzip -v -t ${DB_PATH}/prot.accession2taxid.FULL.gz
printf \"\$(date) - Finished validating 'prot.accession2taxid.FULL.gz'.\n\"
diamond makedb -p \$SLURM_CPUS_PER_TASK --in ${DB_PATH}/${DB}.fasta -d ${DB_PATH}/${DB} --taxonmap ${DB_PATH}/prot.accession2taxid.FULL.gz --taxonnodes ${DB_PATH}/nodes.dmp --taxonnames ${DB_PATH}/names.dmp
printf \"\$(date) - Finished preparing Diamond database '${DB}.dmnd'. \n\"
rm ${DB_PATH}/${DB}.fasta
printf \"\$(date) - Removed '${DB}.fasta', script completed. \n\"" > $JOBNAME.slurm
# send the job to the cluster
sbatch --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CONDA_NAME=genomics $JOBNAME.slurm

WORKDIR=/home/ibar/adna/Oyster_QX_Disease
JOBNAME=prot_diamond_tax
cd $WORKDIR/SRO_TransPi_assembly/Annotation
mkdir -p $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/$JOBNAME
NCORES=12
MEM=96
WALLTIME=50:00:00
ls -1 $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/transdecoder/*.transdecoder.pep | parallel --dry-run --rpl "{sample} s:.+/(.+)\.Trinity.+:\1:" "apptainer run docker://buchfink/diamond:latest blastp -d ~/adna/tools/ncbi_db/nr.dmnd -q {} -p \$SLURM_CPUS_PER_TASK -f 6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle staxids sscinames skingdoms sphylums --sensitive -k 20 -e 0.00001 > $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/$JOBNAME/{sample}.diamond_blastp.nr_tax.outfmt6" > $JOBNAME.cmds


# run within an interactive job
grep "unfixrm_SRO_control" $JOBNAME.cmds | bash
# submit the job array to the cluster
sbatch -a 1-2,4 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=base ${WORKDIR}/array.slurm
# each job took roughly 3:30 hours
```

Using the Nextflow-enabled version of DIAMOND (reducing the time to 1.5h)

```{bash prot-diamond-nf}
start_long_interactive_job # start interactive job on Bunya
WORKDIR=/home/ibar/adna/Oyster_QX_Disease
JOBNAME=nf_diamond_blastp_tax
mkdir -p $WORKDIR/SRO_TransPi_assembly/Annotation/$JOBNAME
cd $WORKDIR/SRO_TransPi_assembly/Annotation/$JOBNAME

# mkdir -p $WORKDIR/Annotation/Annotation_results/$JOBNAME
ls -1 $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/transdecoder/*.transdecoder.pep | parallel --dry-run "~/bin/nextflow-22.11.1-edge-all run ~/adna/tools/nf-blast/nf-blast.nf -profile bunya,apptainer,diamond_tax --query {} --app 'diamond blastp' --db ~/adna/tools/ncbi_db/nr --diamondOpts '--very-sensitive -e 1e-10 -k 20' --chunkSize 10000 --outDir  $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/$JOBNAME --out {/}.dmnd_blastp.nr_tax.outfmt6  -c ~/.nextflow/bunya.config -with-tower" > $JOBNAME.cmds

# run SRO_control interactively
grep "unfixrm_SRO_control" $JOBNAME.cmds | bash
# or submit all the commands as a job array to the cluster

NCORES=4
MEM=24
WALLTIME=250:00:00
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=base ${WORKDIR}/array.slurm

```

###### Functional annotation of predicted proteins with InterPro

The predicted proteins in the transcriptomes were further annotated using InterProScan (or [FA-nf](https://github.com/guigolab/FA-nf)) to assign protein families, motifs and ontologies to assist with transcript-to-gene annotation.
Notice the issues mentioned above with SignalP and TMHMM and a new one for Phobius (see discussions and solutions on [GitHub](https://github.com/nextgenusfs/funannotate/issues/696#issuecomment-1057663190) and [BioStar](https://www.biostars.org/p/238642/))

```{bash prot-annotation}
cd ~/adna/tools
start_interactive_job # start an interactive job to download interproscan
IPSCAN_VERSION=5.66-98.0
IPRO_DATA=interproscan-data-${IPSCAN_VERSION}.tar.gz
# http://ftp.ebi.ac.uk/pub/software/unix/iprscan/5/5.66-98.0/alt/interproscan-data-5.66-98.0.tar.gz
DATA_URL="https://ftp.ebi.ac.uk/pub/software/unix/iprscan/5/${IPSCAN_VERSION}/alt/${IPRO_DATA}"
aria2c -x 6 -c $DATA_URL && \
 aria2c -c $DATA_URL.md5 && \
 md5sum -c ${IPRO_DATA}.md5 && pigz -dc $IPRO_DATA | tar xf -
# download the interproscan container
apptainer pull docker://index.docker.io/interpro/interproscan:${IPSCAN_VERSION} --dir $NXF_APPTAINER_CACHEDIR
# create a link in SINGULARITY_CACHEDIR
ln -s interproscan_${IPSCAN_VERSION}.sif $NXF_APPTAINER_CACHEDIR/

# download the FA-nf pipeline
git clone --recursive https://github.com/guigolab/FA-nf
# copy external software and containers from Research Space
rclone copy -P  GURS_shared:Oyster_QX_Disease/FA-nf/containers FA-nf/containers
cd FA-nf/containers/interproscan && tar xzf external_tools.tar.gz && cd -
# ls -1  FA-nf/containers/interproscan/external/tmhmm-2.0c/bin/* | parallel chmod +x {}
# wget https://github.com/chefnb/ParseTargs/raw/master/tmhmm-2.0c/bin/decodeanhmm -O FA-nf/containers/interproscan/external/phobius/decodeanhmm
# chmod +x FA-nf/containers/interproscan/external/signalp-4.1/signalp FA-nf/containers/interproscan/external/phobius/decodeanhmm  FA-nf/containers/interproscan/external/phobius/phobius.pl

# copy user-defined settings
cp -r FA-nf/containers/interproscan/.interproscan-5 ~/
# edit path to external programs and databases
# nano ~/.interproscan-5/interproscan.properties
( echo "data.directory=$PWD/interproscan-${IPSCAN_VERSION}/data"; sed -r  "s|=.+/FA-nf/containers/interproscan/external|=/home/ibar/scratch/tools|g" $PWD/FA-nf/containers/interproscan/.interproscan-5/interproscan.properties ) > ~/.interproscan-5/interproscan.properties
# sed -i  "s|/home/ibar/etc|$PWD|g" ~/.interproscan-5/interproscan.properties
# echo "data.directory=$PWD/interproscan-${IPSCAN_VERSION}/data" >> ~/.interproscan-5/interproscan.properties
# test that interproscan is working
start_interactive_job # make sure it is run on a compute node
apptainer exec FA-nf/containers/interproscan/interproscan_${IPSCAN_VERSION}.sif /opt/interproscan/interproscan.sh
# press the HMM databases 

apptainer exec -B $PWD/interproscan-${IPSCAN_VERSION}/data:/opt/interproscan/data  $PWD/FA-nf/containers/interproscan/interproscan_${IPSCAN_VERSION}.sif /bin/sh -c "cd /opt/interproscan/; python3 setup.py -f interproscan.properties"
# if the code above isn't working then do it interactively
apptainer shell -B $PWD/interproscan-${IPSCAN_VERSION}/data:/opt/interproscan/data  $PWD/FA-nf/containers/interproscan/interproscan_${IPSCAN_VERSION}.sif
cd /opt/interproscan/; python3 setup.py -f interproscan.properties



# annotate proteins 
WORKDIR=/home/ibar/adna/Oyster_QX_Disease
cd $WORKDIR/SRO_TransPi_assembly/Annotation
mkdir -p $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/ipscan
JOBNAME="SRO_prot_ipscan"
# prepare the commands (don't forget to remove the asterisk at the end of the proteins!)
ls -1 $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/transdecoder/*.transdecoder.pep | parallel --dry-run "sed 's/[*]//g' {} > \$TMPDIR/{/.}.tmp ; apptainer exec -B /home/ibar/scratch/tools -B $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/ipscan:/output -B \$TMPDIR:/temp $NXF_SINGULARITY_CACHEDIR/interproscan_${IPSCAN_VERSION}.sif /opt/interproscan/interproscan.sh -i /temp/{/.}.tmp -d /output -pa -dp -goterms -f TSV -T /temp -cpu \$SLURM_CPUS_PER_TASK ; gawk '\$4~/PANTHER/' $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/ipscan/{/.}.tsv > $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/ipscan/{/.}.panther.tsv" > $JOBNAME.cmds

# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME  ${WORKDIR}/array.slurm

cd $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/ipscan
# remove the tmp from the filenames
rename -v 's/.tmp//' *
# extract PANTHER results
# ls -1 *.tsv | parallel  "gawk '\$4~/PANTHER/' {} > {.}.panther.tsv"
# send results back to Research Space
rclone copy -P $WORKDIR/SRO_TransPi_assembly/Annotation/Annotation_results/ipscan GURS_shared:Oyster_QX_Disease/QX_RNA-Seq/SRO_TransPi_assembly/Annotation_results/ipscan


```

These annotation tables were added to the Trinotate database of the assemblies.

##### Assess the quality of assemblies

The transcriptome assemblies were assessed and compared using [rnaQuast](https://github.com/ablab/rnaquast) v2.2.3 [@bushmanovaRnaQUASTQualityAssessment2016], [TransRate](http://hibberdlab.com/transrate) v1.0.3 [@smith-unnaTransRateReferencefreeQuality2016] and [DETONATE](https://deweylab.biostat.wisc.edu/detonate/) v1.11 [@liEvaluationNovoTranscriptome2014b], as suggested in the Trinity [documentation](https://github.com/trinityrnaseq/trinityrnaseq/wiki/Transcriptome-Assembly-Quality-Assessment) and by @beheraPlantTranscriptomeAssembly2021.

The transcripts were compared to the Pacific Oyster (*Crassostrea gigas*) proteins, which were obtained from NCBI (Accession [GCA_902806645.1](https://www.ncbi.nlm.nih.gov/datasets/genome/GCA_902806645.1/)) [@penalozaChromosomelevelGenomeAssembly2021] 

```{bash eval-assembly}
CONDA_NAME=rnaseq
mamba create -n $CONDA_NAME transrate ncbi-datasets-cli detonate
WORKDIR='/home/ibar/adna/Oyster_QX_Disease'
conda activate $CONDA_NAME

mkdir -p $WORKDIR/references && cd $WORKDIR/references
datasets download genome accession GCF_902806645.1 --include gff3,rna,cds,protein,genome --filename GCF_902806645.1.zip
unzip GCF_902806645.1.zip
REF_PROT=$WORKDIR/references/ncbi_dataset/data/GCF_902806645.1/protein.faa
# $WORKDIR/SRO_TransPi_assembly/results/assemblies
# install the ORP version of transrate
cd ~/etc/tools 
wget https://github.com/macmanes-lab/Oyster_River_Protocol/raw/master/software/orp-transrate.tar.gz
tar xzf orp-transrate.tar.gz

# or install the ORP container
apptainer pull docker://macmaneslab/orp:2.3.3 --dir $NXF_APPTAINER_CACHEDIR

cd $WORKDIR/SRO_TransPi_assembly/

NCORES=24
MEM=96
WALLTIME=100:00:00

#mkdir -p $WORKDIR/SRO_TransPi_assembly/results/transrate
# cp /home/ibar/adna/Oyster_QX_Disease/SRO_TransPi_assembly/work/27/72b07a2131e97971f644da74fc9dc6/unfixrm_SRO_QX_no_rRNA.R1.fq /home/ibar/adna/Oyster_QX_Disease/SRO_TransPi_assembly/work/2b/f20013a2ae897e646817eac5d0997c/SRO_QX_no_rRNA.R1.fq /home/ibar/adna/Oyster_QX_Disease/SRO_TransPi_assembly/work/e5/25ba5d1510380c3631ee6c18bf2e89/SRO_control_no_rRNA.R1.fq /home/ibar/adna/Oyster_QX_Disease/SRO_TransPi_assembly/work/81/5bd2ffe863bf0ee37388694b4ad254/unfixrm_SRO_control_no_rRNA.R1.fq $WORKDIR/SRO_TransPi_assembly/input_files/

# use no_rRNA.R1.fq files as input 
JOBNAME="transrate_no_rRNA"
OUTDIR=$WORKDIR/SRO_TransPi_assembly/results/$JOBNAME # or OUTDIR=~/scratch/sandbox/SRO_TransPi_assembly/$JOBNAME
mkdir -p $OUTDIR
parallel --dry-run --rpl "{sample} s:.+/(.+)_no_rRNA.R1.fq:\1:" --rpl "{read2} s:.R1:.R2:" "mkdir -p $OUTDIR/{sample}; ASSEMBLIES=\$(ls -1 $WORKDIR/SRO_TransPi_assembly/results/assemblies/{sample}*[!k0-9].fa | gawk -v ORS=\",\" '1' | sed 's/,$//'); $HOME/etc/tools/orp-transrate/transrate --assembly \$ASSEMBLIES --left {} --right {read2} --threads \$SLURM_CPUS_PER_TASK --reference $REF_PROT --output $OUTDIR/{sample}" ::: $(ls -1 $WORKDIR/SRO_TransPi_assembly/input_files/*SRO*no_rRNA.R1.fq) > $JOBNAME.cmds
# --merge-assemblies {sample}.merged.fa 
# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME ${WORKDIR}/array.slurm

# use normalised (.norm.fq) files as input 
JOBNAME="transrate_norm"
OUTDIR=$WORKDIR/SRO_TransPi_assembly/results/$JOBNAME # or OUTDIR=$HOME/scratch/sandbox/SRO_TransPi_assembly/$JOBNAME
mkdir -p $OUTDIR
parallel --dry-run --rpl "{sample} s:.+/\w+-(.+).norm.fq:\1:" --rpl "{read2} s:left-:right-:" "mkdir -p $OUTDIR/{sample}; ASSEMBLIES=\$(ls -1 $WORKDIR/SRO_TransPi_assembly/results/assemblies/{sample}*[!k0-9].fa | gawk -v ORS=\",\" '1' | sed 's/,$//'); $HOME/etc/tools/orp-transrate/transrate --assembly \$ASSEMBLIES --left {} --right {read2} --threads \$SLURM_CPUS_PER_TASK --reference $REF_PROT --output $OUTDIR/{sample}" ::: $(find $WORKDIR/SRO_TransPi_assembly/work -type f -name left-*SRO*.norm.fq) > $JOBNAME.cmds
# --merge-assemblies {sample}.merged.fa 
# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME ${WORKDIR}/array.slurm


```

### SRO transcriptome assembly with ORP
The SRO transcriptome of non-infected (control) tissues was also assembled with the Oyster River Protocol pipeline [@macmanesOysterRiverProtocol2018a].

```{bash orp-assembly}
# prepare slurm script
WORKDIR='/scratch/project/adna/Oyster_QX_Disease'
JOBNAME="SRO_orp_pipeline"
NCORES=24
MEM=150
WALLTIME=200:00:00
mkdir -p $WORKDIR/$JOBNAME && cd $WORKDIR/$JOBNAME

echo '#!/bin/bash --login'"
#SBATCH --job-name=$JOBNAME
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log
#SBATCH --cpus-per-task=$NCORES
#SBATCH --mem=${MEM}G
#SBATCH --time=$WALLTIME
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
set -Eeo pipefail
cd \$SLURM_SUBMIT_DIR

apptainer exec -B $WORKDIR:/scratch --no-mount /home $NXF_APPTAINER_CACHEDIR/docker.io-macmaneslab-orp-2.3.3.img bash -c \"source /home/orp/Oyster_River_Protocol/software/anaconda/install/bin/activate orp && unset -f which && cd /scratch/$JOBNAME && /home/orp/Oyster_River_Protocol/oyster.mk TPM_FILT=1 CPU=\$SLURM_CPUS_PER_TASK MEM=$MEM  READ1=/scratch/SRO_TransPi_assembly/input_files/unfixrm_SRO_control_R1.cor.fastq.gz READ2=/scratch/SRO_TransPi_assembly/input_files/unfixrm_SRO_control_R2.cor.fastq.gz RUNOUT=SRO_trans_orp\" " >  $JOBNAME.slurm

# submit the job
sbatch $JOBNAME.slurm


```


#### Gene Expression Analysis

The chosen most representative transcriptome assembly (Rcorrected SRO non-infected assembly) was used as reference to quantify gene/transcript abundance using [Salmon](https://salmon.readthedocs.io/en/latest/salmon.html) [@patroSalmonProvidesFast2017, @loveSwimmingDownstreamStatistical2018] or [Kallisto](https://pachterlab.github.io/kallisto/about) [@brayNearoptimalProbabilisticRNAseq2016]

```{bash salmon-quant}
CONDA_NAME=genomics
mamba install -n $CONDA_NAME kallisto salmon 
WORKDIR=/home/ibar/adna/Oyster_QX_Disease
mkdir -p $WORKDIR/SRO_TransPi_assembly/Gene_expression
cd !$

JOBNAME="Salmon_index"
NCORES=12
MEM=64
WALLTIME=10:00:00
parallel --dry-run "salmon index -t {} -i {/.}.index" ::: $(ls -1 $WORKDIR/SRO_TransPi_assembly/results/assemblies/*Trinity.fa ) > $JOBNAME.cmds 

# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME ${WORKDIR}/array.slurm

DATE=`date +%d_%m_%Y`
JOBNAME="Salmon_quant_$DATE"
RUNDIR=$WORKDIR/SRO_TransPi_assembly/Gene_expression/$JOBNAME
mkdir -p $RUNDIR && cd $RUNDIR
parallel --dry-run --rpl "{sample} s:.+/(.+)_1.fastq.gz:\1:" --rpl "{read2} s:_1:_2:" "salmon quant -i $WORKDIR/SRO_TransPi_assembly/Gene_expression/unfixrm_SRO_control.Trinity.index -l A -1 {} -2 {read2} -o {sample}_salmon_eq --numBootstraps 500 --useVBOpt --seqBias --dumpEq -p \$SLURM_CPUS_PER_TASK --validateMappings" ::: $(ls -1 $WORKDIR/QX_RNA-Seq/data/NN01-5*_1.fastq.gz ) > $JOBNAME.cmds

# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME ${WORKDIR}/array.slurm


```

Expression-based clustering of transcripts was performed with [RapClust](https://github.com/COMBINE-lab/RapClust) [@trapnellDifferentialAnalysisGene2013]. 

```{bash rapclust}
CONDA_NAME=genomics
conda activate $CONDA_NAME
mamba install mcl # need to install mcl
pip install git+https://github.com/IdoBar/RapClust@patch-1
# mamba install -n $CONDA_NAME kallisto salmon rapclust
WORKDIR=/home/ibar/adna/Oyster_QX_Disease
cd $WORKDIR/SRO_TransPi_assembly/Gene_expression/Salmon_quant_14_06_2024

# unzip the eq_classes files
find . -name eq_classes.txt.gz | parallel  "gzip -cd {} > {.}"

JOBNAME="SRO_rapclust"
NCORES=12
MEM=64
WALLTIME=10:00:00
# prepare config file for RapClust
find . -maxdepth 1 -type d -name  "NN01-*salmon_eq" | sort | gawk -F "/" 'BEGIN{printf "conditions:\n  - SRO_Control\n  - SRO_QX\nsamples:\n  SRO_Control:"}{printf "\n    - %s", $2}NR==3{printf "\n  SRO_QX:"}END{printf "\noutdir: SRO_rapclust\n"}' > $JOBNAME.yaml
# prepare SLURM script 
echo "RapClust --config $JOBNAME.yaml" > $JOBNAME.cmds
# submit the job array to the cluster
sbatch -a 1 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=$CONDA_NAME ${WORKDIR}/array.slurm

```


# Genome assembly of _M. sydneyi_

## Analysis Pipeline

### General overview:

1.  Data pre-processing:
    a.  Reads basecalling
    b.  Quality check
    c.  Adaptor trimming
    d.  Post-trim quality check
2.  Mapping reads to host reference genome
3.  _De Novo_ genome assembly of non-host reads
4.  Taxonomic classification of contigs
5.  Removal of additional host and off-target contigs
6.  Annotation of pathogen genes
7.  Summary statistics and visualisation

### Methods

#### DNA Extraction
High molecular DNA was extracted using the Monarch® HMW DNA Extraction Kit for Tissue (New England Biolabs, [NEB T3060](https://www.neb.com/en-au/products/t3060-monarch-hmw-dna-extraction-kit-for-tissue#Product%20Information)) from digestive gland tissue of Sydney rock oysters infected with QX disease. The samples were confirmed to be infected with QX (contain _M. sydneyi_) by a PCR assay using the extracted DNA as template, as described by XXX. DNA integrity, purity and quantity were assessed by gel electrophoresis, Nanodrop and Qubit, respectively.    

#### Nanopore Sequencing and Basecalling
The DNA was prepared for long-read Nanopore sequencing by removal of contaminants (Qiagne [DNeasy PowerClean CleanUp Kit](https://www.qiagen.com/us/products/discovery-and-translational-research/dna-rna-purification/dna-purification/microbial-dna/dneasy-powerclean-cleanup-kit)) and short fragments (ONT Short Fragment Eliminator Kit, EXP-SFE001), followed by ligation of sequencing adapters using the Ligation sequencing DNA V14 (ONT SQK-LSK114) or the Rapid Barcoding kit (SQK-RBK110.96) for sequencing on an Oxford Nanopore Mk1C with a MINFLO114 flowcell (R10) or on a PromethION R9.4.1 flowcell, respectively.   
PromethION data was first demultiplexed with Guppy, then the resulting `fast5` files were converted to `pod5` files using [pod5 package](https://pod5-file-format.readthedocs.io/en/0.2.0/docs/tools.html).  

```{bash setup-conda}
# download and install conda/mamba
wget "https://github.com/conda-forge/miniforge/releases/latest/download/Miniforge3-$(uname)-$(uname -m).sh"
bash Miniforge3-$(uname)-$(uname -m).sh
# initialise conda
source ~/.bashrc
# add channels and set priorities
conda config --add channels conda-forge
conda config --append channels bioconda
# make sure the base environment is always accessible
conda config --set auto_stack 1
# install extra packages to the base environment
conda install -n base libgcc gnutls libuuid readline cmake git tmux libgfortran parallel mamba gawk pigz rename genozip autoconf sshpass pbgzip rclone gh-cli
# setup conda environment for Nanopore data analysis and processing
mamba create -n ont seqkit slow5tools samtools rclone parallel pip numpy pycoqc chopper nanoplot pbgzip
conda activate ont
pip install pod5 pyslow5
```

Basecalling of simplex and duplex reads from the Nanopore sequencing data was performed with [dorado](https://github.com/nanoporetech/dorado).  

The recommended workflow is as follows:  

1. Basecall while aligning to the host reference into `BAM` file (convert `fast5` into `pod5` files if needed)  
2. Generate summary statistics with `dorado summary` and QC (see [this tutorial](https://timkahlke.github.io/LongRead_tutorials/APP_ANS.html#whats-the-name-of-the-configuration-file-guppy-needs-to-basecall-the-tutorial-data))  
3. Save unaligned reads to `FASTQ` format (with `samtools view -Ofastq -f 4 file.bam > unmapped_reads.fastq`, see [this SO thread](https://stackoverflow.com/a/63157775), [this thread](https://bioinformatics.stackexchange.com/a/13616) and [this explanation](https://github.com/samtools/samtools/issues/1153))  
4. Separate duplex reads from simplex to `FASTQ` format (either from the original `BAM` or from the un-mapped `FASTQ`, see [dorado issue #199](https://github.com/nanoporetech/dorado/issues/199#issuecomment-1571617174))  
5. Adaptor trimming of the reads with `Porechop` and quality trimming with `Nanofilt` (average read quality Q>10, minimum read length >500 and removal of the first 10 bases)   


```{bash dorado-setup}
cd ~/aDNA/tools
wget https://cdn.oxfordnanoportal.com/software/analysis/dorado-0.3.4-linux-x64.tar.gz
tar xzfv dorado-0.3.4-linux-x64.tar.gz
ln -s ~/aDNA/tools/dorado-0.3.4-linux-x64/bin/dorado $CONDA_PREFIX/bin/
mkdir dorado_models
cd dorado models
# manually download and unzip dorado models (see  https://github.com/nanoporetech/dorado/issues/266#issuecomment-1613225513)
dorado download --list 2>&1 >/dev/null | grep -v "models" | cut -d " " -f6 | parallel --dry-run 'wget https://cdn.oxfordnanoportal.com/software/analysis/dorado/{}.zip && unzip {}.zip'

```



```{bash dorado-basecall-bam}
WORK_DIR=$HOME/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing

# Niki_QX_1
BATCH=Niki_QX_1_SRO_781
MINQ=7
MODEL="dna_r10.4.1_e8.2_400bps_sup@v4.2.0"
cd $WORK_DIR/Niki_QX_1/SRO_781
mkdir pod5
# create symlinks to pod5 files
find $PWD/202* -name *.pod5 | parallel ln -s {} pod5/

echo "source ~/.bashrc; conda activate ont; module load cuda/12.1; cd $PWD; set -Eeo pipefail; dorado duplex --reference $WORK_DIR/Sro_reference_concatenated.fasta -x cuda:0 --min-qscore $MINQ ~/aDNA/tools/dorado_models/${MODEL} pod5/  > ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam; dorado summary -v ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam > ${BATCH}_sequencing_summary.txt ; samtools view -Ofastq -f 4 ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam | pbgzip -c -8 > ${BATCH}_unmapped_reads.fastq.gz; seqkit grep -nri -p  ';' ${BATCH}_unmapped_reads.fastq.gz -o ${BATCH}_duplex_unmapped_reads_Q7.fastq.gz; seqkit grep -nriv -p  ';' ${BATCH}_unmapped_reads.fastq.gz -o ${BATCH}_simplex_unmapped_reads_Q7.fastq.gz" | qsub -l select=1:ncpus=8:ngpus=1:mem=128gb,walltime=20:00:00 -N Niki_QX_1 -q gpuq2 

# fix it for unmapped reads
ln -s $PWD/Niki_QX_1_SRO_781_reads_Q7.fastq.gz $WORK_DIR/fastq_files/

# Niki_QX_2
BATCH=Niki_QX_2_SRO_781
MINQ=7
MODEL="dna_r10.4.1_e8.2_400bps_sup@v4.2.0"
cd $WORK_DIR/Niki_QX_2/SRO_781
mkdir pod5
# create symlinks to pod5 files
find $PWD/202* -name *.pod5 | parallel ln -s {} pod5/

echo "source ~/.bashrc; conda activate ont; module load cuda/12.1; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/dorado-0.3.1-linux-x64/bin/dorado duplex --reference $WORK_DIR/Sro_reference_concatenated.fasta -x cuda:0 --min-qscore $MINQ ~/aDNA/tools/dorado-0.3.1-linux-x64/dorado_models/${MODEL} pod5/  > ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam; dorado summary -v ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam > ${BATCH}_sequencing_summary.txt ; samtools view -Ofastq -f 4 ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam | pbgzip -c -8 > ${BATCH}_unmapped_reads.fastq.gz; seqkit grep -nri -p  ';' ${BATCH}_unmapped_reads.fastq.gz -o ${BATCH}_duplex_unmapped_reads_Q7.fastq.gz; seqkit grep -nriv -p  ';' ${BATCH}_unmapped_reads.fastq.gz -o ${BATCH}_simplex_unmapped_reads_Q7.fastq.gz" | qsub -l select=1:ncpus=8:ngpus=1:mem=128gb,walltime=20:00:00 -N Niki_QX_2 -q gpuq2

ln -s $PWD/Niki_QX_2_SRO_781_reads_Q7.fastq.gz $WORK_DIR/fastq_files/

# separate duplex from simplex reads
seqkit grep -nri -p  ";" Niki_QX_2_raw_reads_Q7.fastq -o Niki_QX_2_duplex_reads_Q7.fastq.gz
seqkit grep -nriv -p  ";" Niki_QX_2_raw_reads_Q7.fastq -o Niki_QX_2_simplex_reads_Q7.fastq.gz
# calculate percent of duplex reads
DUPS=$(zcat Niki_QX_2_duplex_reads_Q7.fastq.gz | gawk 'NR%4==1' | wc -l)
TOTAL=$(zcat Niki_QX_2_SRO_781_reads_Q7.fastq.gz | gawk 'NR%4==1' | wc -l)
DUP_RATIO=$(bc <<<"scale=4;$DUPS/$TOTAL")


# process the PromethION batch
BATCH=Niki_QX_SRO_781_Prom1
PROM_DIR=$HOME/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing/PromethION_QAAFI/20230612_1541_1G_PAG73069_dca304f3
MODEL="dna_r9.4.1_e8_sup@v3.6"
mkdir -p $PROM_DIR/fast5_pass

# copy data from Research Space
rclone copy -P GU_owncloud:Shared/Oyster_QX_Disease/Nanopore_sequencing/A_rabiei_M_sydneyi/20230612_1541_1G_PAG73069_dca304f3/fast5_pass/barcode19 $PROM_DIR/fast5_pass/barcode19
rclone copy -P GU_owncloud:Shared/Oyster_QX_Disease/Nanopore_sequencing/A_rabiei_M_sydneyi/20230612_1541_1G_PAG73069_dca304f3/fast5_pass/barcode20 $PROM_DIR/fast5_pass/barcode20

cd $PROM_DIR/fast5_pass
# convert from fast5 to pod5
pod5 convert fast5 ./barcode19 --output ./barcode19/pod5/Niki_QX_QAAFI_barcode19.pod5
pod5 convert fast5 ./barcode20 --output ./barcode20/pod5/Niki_QX_QAAFI_barcode20.pod5



#ls -1 barcode* | parallel --dry-run "module load cuda/12.1; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/dorado-0.3.1-linux-x64/bin/dorado basecaller -r -x cuda:0 --emit-fastq --min-qscore 9 ~/aDNA/tools/dorado-0.3.1-linux-x64/dorado_models/dna_r9.4.1_e8_sup@v3.6 {} > Niki_QX_QAAFI_{}_raw_reads_Q9.fastq" | qsub -l select=1:ncpus=4:ngpus=1:mem=16gb,walltime=20:00:00 -N Niki_QX_QAAFI -q gpuq2

# create symlinks to pod5 files
# find $PWD/202* -name *.pod5 | parallel ln -s {} pod5/

echo "source ~/.bashrc; conda activate ont; module load cuda/12.1; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/dorado-0.3.1-linux-x64/bin/dorado basecaller --reference $WORK_DIR/Sro_reference_concatenated.fasta -x cuda:0 --min-qscore $MINQ ~/aDNA/tools/dorado-0.3.1-linux-x64/dorado_models/$MODEL barcode20/pod5/  > ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam; dorado summary -v ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam > ${BATCH}_sequencing_summary.txt ; samtools view -Ofastq -f 4 ${BATCH}_SRO-mapped_reads_Q${MINQ}.bam | pbgzip -c -8 > ${BATCH}_unmapped_reads.fastq.gz" | qsub -l select=1:ncpus=8:ngpus=1:mem=128gb,walltime=20:00:00 -N Niki_QX_Prom1 -q gpuq2

ln -s $PWD/Niki_QX_2_SRO_781_reads_Q7.fastq.gz $WORK_DIR/fastq_files/


# copy the files to Research Space
rclone copy -P -L  $WORK_DIR/fastq_files GU_owncloud:Shared/Oyster_QX_Disease/Nanopore_sequencing/fastq_files

```

<!-- ```{bash dorado-basecall-fastq} -->


<!-- # run interactively (Niki_QX_1): -->
<!-- qsub -I -l select=1:ncpus=4:ngpus=1:mem=12gb,walltime=10:00:00 -q gpuq2 -N ONT-dorado -->

<!-- module load cuda/12.1 -->
<!-- # export LD_LIBRARY_PATH=/scratch/aDNA/tools/dorado-0.3.1-linux-x64/lib:$LD_LIBRARY_PATH -->

<!-- cd ~/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing/Niki_QX_1/SRO_781/20230619_1431_MC-114875_FAW74280_8dcf3c8a/ -->

<!-- # run basecalling in Duplex mode -->
<!-- ~/aDNA/tools/dorado-0.3.1-linux-x64/bin/dorado duplex -x cuda:0 --emit-fastq --min-qscore 9 ~/aDNA/tools/dorado-0.3.1-linux-x64/dorado_models/dna_r10.4.1_e8.2_400bps_sup@v4.2.0 pod5/ > Niki_QX_1_raw_reads_Q9.fastq -->

<!-- # or submit to the cluster (Niki_QX_2) -->

<!-- cd ~/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing/Niki_QX_2/SRO_781 -->
<!-- mkdir pod5 -->
<!-- # create symlinks to pod5 files -->
<!-- find $PWD/202* -name *.pod5 | parallel ln -s {} pod5/ -->

<!-- echo "source ~/.bashrc; conda activate ont; module load cuda/12.1; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/dorado-0.3.1-linux-x64/bin/dorado duplex -r -x cuda:0 --emit-fastq --min-qscore 7 ~/aDNA/tools/dorado-0.3.1-linux-x64/dorado_models/dna_r10.4.1_e8.2_400bps_sup@v4.2.0 pod5/ | pbgzip -6 -c > Niki_QX_2_raw_reads_Q7.fastq.gz; seqkit grep -nri -p  ';' Niki_QX_2_raw_reads_Q7.fastq.gz -o Niki_QX_2_duplex_reads_Q7.fastq.gz; seqkit grep -nriv -p  ';' Niki_QX_2_raw_reads_Q7.fastq.gz -o Niki_QX_2_simplex_reads_Q7.fastq.gz" | qsub -l select=1:ncpus=4:ngpus=1:mem=16gb,walltime=20:00:00 -N Niki_QX_2 -q gpuq2 -->

<!-- # separate duplex from simplex reads -->
<!-- seqkit grep -nri -p  ";" Niki_QX_2_raw_reads_Q7.fastq -o Niki_QX_2_duplex_reads_Q7.fastq.gz -->
<!-- seqkit grep -nriv -p  ";" Niki_QX_2_raw_reads_Q7.fastq -o Niki_QX_2_simplex_reads_Q7.fastq.gz -->
<!-- # calculate percent of duplex reads -->
<!-- DUPS=$(zcat Niki_QX_2_duplex_reads_Q7.fastq.gz | gawk 'NR%4==1' | wc -l) -->
<!-- TOTAL=$(gawk 'NR%4==1' Niki_QX_2_raw_reads_Q7.fastq | wc -l ) -->
<!-- DUP_RATIO=$(bc <<<"scale=4;$DUPS/$TOTAL") -->

<!-- # process the PromethION batch -->
<!-- PROM_DIR=$HOME/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing/PromethION_QAAFI/20230612_1541_1G_PAG73069_dca304f3 -->
<!-- mkdir -p $PROM_DIR/fast5_pass -->

<!-- # copy data from Research Space -->
<!-- rclone copy -P GU_owncloud:Shared/Oyster_QX_Disease/Nanopore_sequencing/A_rabiei_M_sydneyi/20230612_1541_1G_PAG73069_dca304f3/fast5_pass/barcode19 $PROM_DIR/fast5_pass/barcode19 -->
<!-- rclone copy -P GU_owncloud:Shared/Oyster_QX_Disease/Nanopore_sequencing/A_rabiei_M_sydneyi/20230612_1541_1G_PAG73069_dca304f3/fast5_pass/barcode20 $PROM_DIR/fast5_pass/barcode20 -->

<!-- cd $PROM_DIR/fast5_pass -->
<!-- # convert from fast5 to pod5 -->
<!-- pod5 convert fast5 ./barcode19 --output Niki_QX_QAAFI_barcode19.pod5 -->
<!-- pod5 convert fast5 ./barcode20 --output Niki_QX_QAAFI_barcode20.pod5 -->

<!-- ls -1 barcode* | parallel --dry-run "module load cuda/12.1; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/dorado-0.3.1-linux-x64/bin/dorado basecaller -r -x cuda:0 --emit-fastq --min-qscore 9 ~/aDNA/tools/dorado-0.3.1-linux-x64/dorado_models/dna_r9.4.1_e8_sup@v3.6 {} > Niki_QX_QAAFI_{}_raw_reads_Q9.fastq" | qsub -l select=1:ncpus=4:ngpus=1:mem=16gb,walltime=20:00:00 -N Niki_QX_QAAFI -q gpuq2 -->

<!-- # separate duplex from simplex reads -->
<!-- seqkit grep -nri -p  ";" Niki_QX_1_raw_reads_Q9.fastq -o Niki_QX_1_duplex_reads_Q9.fastq.gz -->
<!-- seqkit grep -nriv -p  ";" Niki_QX_1_raw_reads_Q9.fastq -o Niki_QX_1_simplex_reads_Q9.fastq.gz -->
<!-- # calculate percent of duplex reads -->
<!-- DUPS=$(zcat Niki_QX_1_duplex_reads_Q9.fastq.gz | gawk 'NR%4==1' | wc -l) -->
<!-- TOTAL=$(gawk 'NR%4==1' Niki_QX_1_raw_reads_Q9.fastq | wc -l ) -->
<!-- DUP_RATIO=$(bc <<<"scale=4;$DUPS/$TOTAL") -->

<!-- ``` -->

#### Non-host Assembly

The next steps are to upload the raw reads to the Griffith University Research Space and assemble the non-host reads with `Flye` [@kolmogorovAssemblyLongErrorprone2019, @kolmogorovMetaFlyeScalableLongread2020] (using `--meta` and `--nano-hq` flags and toggle the 'Enable scaffolding using graph' option in [GalaxyAU](https://usegalaxy.org.au)).  
_Need to find guides/papers about assembling metagenomes with Nanopore data_ -- We need to find which tools/parameters to use. See additional information in [@benoitEfficientHighQualityMetagenome2023, @fengMetagenomeAssemblyHighfidelity2022, @arumugamRecoveryAnalysisLongRead2023]  
In addition, it would be great to have a reference genome of a closely related species (see [list of genomes](https://www.ncbi.nlm.nih.gov/genome/?term=Rhizaria) on) that can be used as a guide for [scaffolding using `RagTag`](https://github.com/malonge/RagTag/wiki/scaffold).  

```{bash flye-assembly}


```

##### Comparison to Ascetosporean genomes

The assembled contigs were aligned to the newly published genomes of several Ascetosporea protists (including _Marteilia pararefringens_, see [NCBI Bioproject PRJNA1051017](https://www.ncbi.nlm.nih.gov/bioproject/PRJNA1051017); @hiltunenthorenComparativeGenomicsAscetosporea2024) using [nf-core/pairgenomealign](https://nf-co.re/pairgenomealign/1.1.0/) v1.1.0 and `mummer` [@marcaisMUMmer4FastVersatile2018].  

```{bash pairgenomealign}
# download Flye assembly from research space.
mkdir -p $HOME/adna/Oyster_QX_Disease/Nanopore_sequencing/SRO_781
rclone copy -P GURS_shared:Oyster_QX_Disease/Nanopore_sequencing/SRO_781/Galaxy_analysis Galaxy_analysis

# WORK_DIR=$HOME/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing # on Gowonda
WORK_DIR=$HOME/adna/Oyster_QX_Disease/Comparative_Genomics # on Bunya
mkdir -p $WORK_DIR && cd $WORK_DIR

# download Ascetosporea genomes from NCBI
conda activate genomics
datasets download genome accession PRJNA1051017 --filename Ascetosporea.zip
# download metadata
# datasets summary genome accession PRJNA1051017  | jq -r '.genes.genes[].gene | [.gene_id, .common_name, .taxname, .tax_id, .transcripts[0].accession_version, .chromosomes[0]] | @csv' > primates_mc1r.csv
# prepare sample sheet
echo "sample,fasta" > samplesheet.csv
find $PWD/ncbi_dataset/data -name "*genomic.fna" | parallel --rpl "{query} s:.+/GCA_.+\.1_UU_(.+)_1_genomic.fna:\1:" echo "{query},{}" >> samplesheet.csv

start_long_interactive_job
TARGET="$HOME/adna/Oyster_QX_Disease/Nanopore_sequencing/SRO_781/Galaxy_analysis/Galaxy_Flye_assembly.fasta"
~/bin/nextflow-24.04.4-all run nf-core/pairgenomealign  -c ~/.nextflow/bunya.config -profile bunya,apptainer  --target $TARGET  --input samplesheet.csv --outdir Ascetosporea_pairedgenomealign

rclone copy -P Ascetosporea_pairedgenomealign GURS_shared:Oyster_QX_Disease/Comparative_Genomics/Ascetosporea_pairedgenomealign
```



```{bash mummer-genomealign}
CONDA_NAME="genomics"
mamba install -n $CONDA_NAME mummer4

WORK_DIR=$HOME/adna/Oyster_QX_Disease/Comparative_Genomics # on Bunya
mkdir -p $WORK_DIR && cd $WORK_DIR
QUERY="$HOME/adna/Oyster_QX_Disease/Nanopore_sequencing/SRO_781/Galaxy_analysis/Galaxy_Flye_assembly.fasta"
JOB_NAME="Ascetosporea_mummer"

find $WORK_DIR/ncbi_dataset/data -name "*genomic.fna" | parallel --dry-run --rpl "{target} s:.+/GCA_.+\.1_UU_(.+)_1_genomic.fna:\1:" nucmer -t $SLURM_CPUS_PER_TASK -p Msyd_{target}_nucmer $QUERY {} > $JOB_NAME.cmds 


```

The contig **matches** between the 2 genomes (synteny plots), were plotted with [SYNY](https://github.com/PombertLab/SYNY) v1.2 [@julianSYNYPipelineInvestigate2024].

```{bash syny}
CONDA_NAME="syny"
mamba create -n $CONDA_NAME syny perl-gd
ln -s /home/ibar/miniconda3/envs/syny/lib/perl5/site_perl/5.22.0/x86_64-linux/GD $CONDA_PREFIX/lib/perl5/site_perl/
conda activate syny
NCORES=8
MEM=16
WALLTIME=1:00:00
WORKDIR=$HOME/adna/Oyster_QX_Disease/Comparative_Genomics # on Bunya
mkdir -p $WORKDIR && cd $WORKDIR
QUERY="$HOME/adna/Oyster_QX_Disease/Nanopore_sequencing/SRO_781/Galaxy_analysis/Galaxy_Flye_assembly.fasta"
fasta_to_gbff.pl -v -f $QUERY -g -o ./
# run SYNY
JOB_NAME="Ascetosporea_SYNY"
find $WORKDIR/ncbi_dataset/data -name "*genomic.fna" | parallel --dry-run --rpl "{target} s:.+/GCA_.+\.1_UU_(.+)_1_genomic.fna:\1:" "fasta_to_gbff.pl -v -f {} -g -o Msyd_{target}_SYNY; ln -s $WORKDIR/Galaxy_Flye_assembly.gbff.gz Msyd_{target}_SYNY/; run_syny.pl -a Msyd_{target}_SYNY/*.gbff.gz -o Msyd_{target}_SYNY -t \$SLURM_CPUS_PER_TASK -e 1e-10" > $JOB_NAME.cmds
 
# submit the job array to the cluster
sbatch -a 1-$(cat $JOB_NAME.cmds | wc -l) --job-name=$JOB_NAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,CMDS_FILE=$JOB_NAME.cmds,CONDA_NAME=$CONDA_NAME ${WORKDIR}/../array.slurm

# focus on matching contigs only (after alignment of the genomes with Minimap2 within SYNY)
start_interactive_job
conda activate syny
WORKDIR=$HOME/adna/Oyster_QX_Disease/Comparative_Genomics # on Bunya
mkdir -p $WORKDIR && cd $WORKDIR
cut -f1 Msyd_MPar_SYNY/ALIGNMENTS/PAF/GCA_039583845_vs_Galaxy_Flye_assembly.mmap.paf | uniq > Msyd_Mpar_shared_contigs.txt
cut -f6 Msyd_MPar_SYNY/ALIGNMENTS/PAF/GCA_039583845_vs_Galaxy_Flye_assembly.mmap.paf | uniq >> Msyd_Mpar_shared_contigs.txt

run_syny.pl -a Msyd_MPar_SYNY/*.gbff.gz -o Msyd_MPar_subset_SYNY --include Msyd_Mpar_shared_contigs.txt -t $SLURM_CPUS_PER_TASK -e 1e-10

```

##### Contig annotation and taxonomic classification

The assembled contigs can then be annotated with `BLASTn` for taxonomic classification.

```{bash tax-class}
# update blast databases
conda activate snippy
source ~/.proxy
mamba install google-cloud-sdk awscli
cd /project/aerc/ncbi/v5
update_blastdb.pl --decompress --source gcp --num_threads 12 ref_euk_rep_genomes swissprot nr refseq_protein # taxdb nt refseq_rna ref_prok_rep_genomes  
# run blast
WORK_DIR=$HOME/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing
cd $WORK_DIR
QUERY="Galaxy_Flye_assembly"
echo "source ~/.bashrc; conda activate snippy; cd $PWD; set -Eeo pipefail; blastn -task blastn -db nt -query $QUERY.fasta -outfmt '6 std stitle staxids ssciname scomname' -evalue 1e-10 -max_target_seqs 20 -num_threads \$NCPUS -out $QUERY.nt_tax.outfmt6" | qsub -l select=1:ncpus=12:mem=64gb,walltime=150:00:00 -N Ass_blast -q longbigmem
```

However, due to the local alignment nature of BLAST, it can result in multiple hits per contig that are likely to disagree, so a better approach would be to classify all the proteins/genes predicted within a contig and determine the most likely organism (or Last Common Ancestor - LCA). This approach for taxonomic classification of the assembled contigs is implemented in a pipeline combining `DIAMOND` and `MEGAN`, and was performed according to [@bagciDIAMONDMEGANFastEasy2021].  
*Note that the performance of `DIAMOND` can be improved when running on a large memory node on an HPC as follows:*

> on a compute server with high amount of RAM, it is useful to set `-b` to a higher number and `-c` to a lower number (e.g., `-b 12 -c 1`). Additionally, the parameter `-t` sets the directory where DIAMOND writes temporary files. If the server on which `DIAMOND` is run has a virtual filesystem, such as `TMPDIR`, setting the temporary directory for `DIAMOND` to this filesystem will also make it perform significantly better, as the IO operations will not be limited by slow network filesystems (e.g., `-t $TMPDIR`). Please see http://www.diamondsearch.org for more details.

```{bash tax-class-dmnd}
# prepare databases
DB_DIR="/project/aerc"
mkdir -p $DB_DIR/megan
cd $DB_DIR/megan
source ~/.proxy
conda activate base
mamba install aria2 ncbi-datasets-cli
# download the nucleotide database
aria2c --header 'Host: software-ab.cs.uni-tuebingen.de' --user-agent 'Mozilla/5.0 (Windows NT 10.0; Win64; x64; rv:109.0) Gecko/20100101 Firefox/117.0' --header 'Accept: text/html,application/xhtml+xml,application/xml;q=0.9,image/avif,image/webp,*/*;q=0.8' --header 'Accept-Language: en-AU,en-US;q=0.7,en;q=0.3' --header 'DNT: 1' --referer 'https://software-ab.cs.uni-tuebingen.de/download/megan6/welcome.html' --header 'Upgrade-Insecure-Requests: 1' --header 'Sec-Fetch-Dest: document' --header 'Sec-Fetch-Mode: navigate' --header 'Sec-Fetch-Site: same-origin' --header 'Sec-Fetch-User: ?1' --header 'Sec-GPC: 1' 'https://software-ab.cs.uni-tuebingen.de/download/megan6/megan-nucl-Feb2022.db.zip' --out 'megan-nucl-Feb2022.db.zip' && unzip megan-nucl-Feb2022.db.zip

# download protein database
aria2c --header 'Host: software-ab.cs.uni-tuebingen.de' --user-agent 'Mozilla/5.0 (Windows NT 10.0; Win64; x64; rv:109.0) Gecko/20100101 Firefox/117.0' --header 'Accept: text/html,application/xhtml+xml,application/xml;q=0.9,image/avif,image/webp,*/*;q=0.8' --header 'Accept-Language: en-AU,en-US;q=0.7,en;q=0.3' --header 'DNT: 1' --referer 'https://software-ab.cs.uni-tuebingen.de/download/megan6/welcome.html' --header 'Upgrade-Insecure-Requests: 1' --header 'Sec-Fetch-Dest: document' --header 'Sec-Fetch-Mode: navigate' --header 'Sec-Fetch-Site: same-origin' --header 'Sec-Fetch-User: ?1' --header 'Sec-GPC: 1' 'https://software-ab.cs.uni-tuebingen.de/download/megan6/megan-map-Feb2022.db.zip' --out 'megan-map-Feb2022.db.zip' && unzip megan-map-Feb2022.db.zip


# prepare the nr as a DIAMOND database - this requires the precompiled binary from github (doesn't work with the conda version yet)
source ~/.proxy
# conda activate ont
# mamba install diamond # megan
# install MEGAN manually
MEGAN_VER="6_25_3"
cd ~/aDNA/tools
aria2c https://software-ab.cs.uni-tuebingen.de/download/megan6/MEGAN_Community_unix_${MEGAN_VER}.sh
bash MEGAN_Community_unix_${MEGAN_VER}.sh -c # choose to install in ~/aDNA/tools/megan
# download a diamond version that can work on NCBI databases
aria2c http://github.com/bbuchfink/diamond/releases/download/v2.1.8/diamond-linux64.tar.gz
tar xzf diamond-linux64.tar.gz

DB_DIR="/project/aerc"
WORK_DIR=$HOME/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing
cd $WORK_DIR

QUERY="Galaxy_Flye_assembly"
CONDA_ENV="ont"

PREP_DB=$(echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/diamond prepdb -d $DB_DIR/ncbi/v5/nr --threads \$NCPUS" | qsub -l select=1:ncpus=8:mem=64gb,walltime=5:00:00 -N prep_db -q sparks2 |  egrep -o "^[0-9]+")  # this takes roughly 1 hour

# run diamond blastx
DMND_BLASTX=$(echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/diamond blastx --sensitive -q $QUERY.fasta -d $DB_DIR/ncbi/v5/nr -o $QUERY.daa  -f 100 --long-reads -p \$NCPUS -t \$TMPDIR -b 10 -c 2" | qsub -l select=1:ncpus=12:mem=96gb,walltime=10:00:00 -N dmnd_blastx -q longbigmem -W depend=afterok:$PREP_DB |  egrep -o "^[0-9]+") # this takes roughly 2 hours (see comment in the markdown)
# prepare database to run DIAMOND with Taxonomy
PREP_DB_TAX=$(echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/diamond makedb --in $DB_DIR/ncbi/v5/nr.gz -d $DB_DIR/ncbi/v5/nr.dmnd --threads \$NCPUS --taxonmap $DB_DIR/ncbi/v5/prot.accession2taxid.FULL.gz --taxonnames $DB_DIR/ncbi/v5/names.dmp --taxonnodes $DB_DIR/ncbi/v5/nodes.dmp" | qsub -l select=1:ncpus=8:mem=64gb,walltime=5:00:00 -N prep_db -q sparks2 |  egrep -o "^[0-9]+")
# run DIAMOND blastx (with taxonomy)
DMND_BLASTX_TAX=$(echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/diamond blastx --sensitive -q $QUERY.fasta -d $DB_DIR/ncbi/v5/nr.dmnd --outfmt 6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle staxids sscinames sphylums -o $QUERY.nr_tax.outfmt6.txt --long-reads -p \$NCPUS -t \$TMPDIR -b 10 -c 2" | qsub -l select=1:ncpus=12:mem=96gb,walltime=50:00:00 -N dmnd_blastx_tax -q longbigmem  |  egrep -o "^[0-9]+")

# run DAA-meganizer - need version >6.25.1 (see https://megan.cs.uni-tuebingen.de/t/meganizing-stuck-at-writing-80/2121/10)

printf "Mollusca\nBacteria\nViruses\n" > contaminants.txt
MEGANIZER=$(echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; xvfb-run --auto-servernum --server-num=1 ~/aDNA/tools/megan/tools/daa-meganizer -i $PWD/$QUERY.daa -mdb $DB_DIR/megan/megan-map-Feb2022.db -cf $PWD/contaminants.txt -t \$NCPUS" | qsub -l select=1:ncpus=12:mem=64gb,walltime=150:00:00 -N meganizer -q longbigmem -W depend=afterok:$DMND_BLASTX |  egrep -o "^[0-9]+") # this shouldn't long time - maybe it finished with 1GB and 1 minute?

# if daa-meganizer is not working then we can use daa2rma
DAA2RMA=$(echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/megan/tools/daa2rma --in $QUERY.daa -mdb $DB_DIR/megan/megan-map-Feb2022.db -cf contaminants.txt -lg true -v true -t \$NCPUS -o $QUERY.rma" | qsub -l select=1:ncpus=12:mem=64gb,walltime=150:00:00 -N daa2rma -q longbigmem |  egrep -o "^[0-9]+")

# convert DIAMOND results to a tsv BLAST-like output
DAA2TSV=$(echo "source ~/.bashrc; cd $PWD; set -Eeo pipefail; ~/aDNA/tools/diamond view -a $QUERY.daa --threads \$NCPUS --out $QUERY.outfmt6.tsv --outfmt 6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore stitle" | qsub -l select=1:ncpus=12:mem=64gb,walltime=10:00:00 -N daa2tsv -q bigmem |  egrep -o "^[0-9]+")

cut -f2,2 $QUERY.outfmt6.tsv | parallel "datasets summary gene accession {} --as-json-lines | dataformat tsv gene --elide-header --fields description,tax-name,tax-id,name-id | paste <(echo {}) - >> Galaxy_Flye_assembly.diamond_nr.annot.outfmt6.tsv"

```

The DIAMOND results were converted to a `tsv` format (`--outfmt 6`) and the accession numbers and the matching titles/descriptions and taxonomy information with retrieved with `rentrez` and `taxonomizr`

```{r process-diamond}
dmnd_headers <- c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", 
                   "qstart", "qend", "sstart", "send", "evalue", "bitscore",
                   "stitle", 'staxids', 'sscinames', 'sphylums')
diamond_res <- read_tsv("data/Galaxy_Flye_assembly.nr_tax.outfmt6.txt", 
                        col_names = dmnd_headers) 
best_diamond_res <- diamond_res %>% group_by(qseqid) %>% arrange(desc(bitscore)) %>% 
  slice(1) %>% write_csv("results/Galaxy_Flye_assembly.diamond_nr.best_hits.csv")

# taxonomizr::prepareDatabase(sqlFile = "data/NCBI_accession2taxa.sql", getAccessions = TRUE)
diamond_tax_results <- accessionToTaxa(diamond_res$stitle[1], "data/NCBI_accession2taxa.sql")

```

##### Taxonomic classification and assembly summary with Blobtoolkit

An alternative approach is to use 'SprayNPray', as detailed in [@garberSprayNPrayUserfriendlyTaxonomic2022] or BlobToolKit ([BlobTools2](https://blobtoolkit.genomehubs.org/blobtools2/blobtools2-tutorials/getting-started-with-blobtools2)) [@challisBlobToolKitInteractiveQuality2020]

1. Create a Blobtoolkit dataset ([link](https://blobtoolkit.genomehubs.org/blobtools2/blobtools2-tutorials/creating-a-dataset/))  
2. Add BLAST and Diamond results ([link](https://blobtoolkit.genomehubs.org/blobtools2/blobtools2-tutorials/getting-started-with-blobtools2/#add_blast))  
3. Add coverage information ([link](https://blobtoolkit.genomehubs.org/blobtools2/blobtools2-tutorials/adding-data-to-a-dataset/adding-coverage/))  
4. Add BUSCO results ([link](https://blobtoolkit.genomehubs.org/blobtools2/blobtools2-tutorials/adding-data-to-a-dataset/adding-busco/))  
3. Visualise the results ([link](https://blobtoolkit.genomehubs.org/btk-viewer/viewer-tutorials/))

```{bash blobtoolkit}
# install blobtoolkit
CONDA_NAME=blobtk
mamba create -n $CONDA_NAME firefox geckodriver
conda activate $CONDA_NAME
pip install "blobtoolkit[full]"

# download taxonomy databases
TX_DB_DIR="/project/aerc/ncbi/new_taxdump"
mkdir -p $TX_DB_DIR
cd $TX_DB_DIR
curl -L https://ftp.ncbi.nih.gov/pub/taxonomy/new_taxdump/new_taxdump.tar.gz | tar xzf -

# setup working directory
WORK_DIR=$HOME/scratch/McDougall/Niki_QX_genome/Nanopore_sequencing
cd $WORK_DIR
QUERY="Galaxy_Flye_assembly"


# create dataset
echo "assembly:
  accession: GCA_001028725.1
  alias: S_venezuelensis_HH1
  bioproject: PRJEB530
  biosample: AMD00012916
  record_type: contig
taxon:
  name: Martylia sydneyi" > $QUERY.yaml

echo "source ~/.bashrc; cd $PWD; conda activate $CONDA_NAME; set -Eeo pipefail; blobtools create --fasta $QUERY.fasta --taxid 98371 --taxdump $TX_DB_DIR $QUERY" | qsub -l select=1:ncpus=8:mem=32gb,walltime=10:00:00 -N  # --meta $QUERY.yaml \

# add diamond and BLAST hits
echo "source ~/.bashrc; cd $PWD; conda activate $CONDA_NAME; set -Eeo pipefail; blobtools add --bitscore 100 --hits $QUERY.nt_tax.outfmt6 --hits $QUERY.nr_tax.outfmt6.txt --taxrule bestsumorder --taxdump $TX_DB_DIR --hits-cols 1=qseqid,14=staxids,12=bitscore,2=sseqid,7=qstart,8=qend,11=evalue $QUERY" | qsub -l select=1:ncpus=8:mem=32gb,walltime=10:00:00 -N add_blob_hits 
    
# open an interactive viewer
qsub -I -X -l select=1:ncpus=8:mem=32gb,walltime=10:00:00 -N view_blobtk 
conda activate blobtk
blobtools view --interactive $QUERY
blobtools host 

```

##### Taxonomic classification with MMseq2

```{bash mmseq-tax-class}

```

#### Gene annotation

Gene models were predicted using a combination of _ab-initio_ and evidence-based (homology to known proteins and transcripts from close species) approaches. These were implemented in the [nf-core/genomeannotator](https://nf-co.re/genomeannotator) and [funannotate](https://funannotate.readthedocs.io/en/latest/tutorials.html) pipelines.


```{bash prep-qx-trans}
CONDA_NAME=genomics
mamba install -n $CONDA_NAME how_are_we_stranded_here
# Prepare QX transcriptome to train gene models
RNASEQ_DIR="/scratch/project/adna/Oyster_QX_Disease/QX_RNA-Seq"
# download QX Trinity transcriptome assembly from Research Space
cd $RNASEQ_DIR
rclone copyto -P GURS_shared:Oyster_QX_Disease/QX_RNA-Seq/QX_TransPi_assembly/assemblies/NN01-QX_R.Trinity.fa $RNASEQ_DIR/NN01-QX_R.Trinity.fa
# download RNA-seq reads
rclone copy -P GURS_shared:Oyster_QX_Disease/QX_RNA-Seq/data/unaligned_reads $RNASEQ_DIR/data/unaligned_reads
# Make a GTF file from transcriptome
$(find $HOME/etc/tools -name "cdna_fasta_file_to_transcript_gtf.pl") $RNASEQ_DIR/NN01-QX_R.Trinity.fa > $RNASEQ_DIR/NN01-QX_R.Trinity.gtf
# clean headers in assembly
reformat.sh in=NN01-QX_R.Trinity.fa out=NN01-QX_R.Trinity.trimmed_headers.fa trd=t
# check strandness with how_are_we_stranded_here script 
check_strandedness --gtf $RNASEQ_DIR/NN01-QX_R.Trinity.gtf --transcripts $HOME/adna/Oyster_QX_Disease/QX_RNA-Seq/NN01-QX_R.Trinity.trimmed_headers.fa --reads_1 $RNASEQ_DIR/data/unaligned_reads/NN01-567_unaligned_reads_R1.fastq.gz --reads_2 $RNASEQ_DIR/data/unaligned_reads/NN01-567_unaligned_reads_R2.fastq.gz


# check strandness with script from ORP (must be run within an interctive job on a compute node)
apptainer exec -B $RNASEQ_DIR:/scratch --no-mount /home $NXF_APPTAINER_CACHEDIR/docker.io-macmaneslab-orp-2.3.3.img bash -c "source /home/orp/Oyster_River_Protocol/software/anaconda/install/bin/activate orp_trinity && unset -f which && cd /scratch && sed -r 's|MAKEDIR :=.+|MAKEDIR := /home/orp/Oyster_River_Protocol|; s|MAKEDIR}version|MAKEDIR}/version|' /home/orp/Oyster_River_Protocol/strandeval.mk > ./strandeval.mk && chmod +x ./strandeval.mk &&  ./strandeval.mk main ASSEMBLY=./NN01-QX_R.Trinity.trimmed_headers.fa CPU=$SLURM_CPUS_PER_TASK READ1=./data/unaligned_reads/NN01-567_unaligned_reads_R1.fastq.gz READ2=./data/unaligned_reads/NN01-567_unaligned_reads_R2.fastq.gz RUNOUT=NN01-QX" 

# check strandess of SRO assembly
JOBNAME="check-strandness-SRO-TransPi"
NCORES=8
MEM=24
WALLTIME=2:00:00
SRO_RNASEQ_DIR="/scratch/project/adna/Oyster_QX_Disease/SRO_TransPi_assembly"
cd $SRO_RNASEQ_DIR
# copy fastq files from Research Space
rclone copyto -P GURS_shared:Oyster_QX_Disease/QX_RNA-Seq/data/corrected_reads/unfixrm_SRO_control_R1.cor.fastq.gz $SRO_RNASEQ_DIR/input_files/unfixrm_SRO_control_R1.cor.fastq.gz
rclone copyto -P GURS_shared:Oyster_QX_Disease/QX_RNA-Seq/data/corrected_reads/unfixrm_SRO_control_R2.cor.fastq.gz $SRO_RNASEQ_DIR/input_files/unfixrm_SRO_control_R2.cor.fastq.gz


find ./results/assemblies -name "*.Trinity.fa" | parallel --dry-run --rpl "{ass} s:.+/(.+).Trinity.fa:\1:" "apptainer exec -B $SRO_RNASEQ_DIR:/scratch --no-mount /home $NXF_APPTAINER_CACHEDIR/docker.io-macmaneslab-orp-2.3.3.img bash -c \"source /home/orp/Oyster_River_Protocol/software/anaconda/install/bin/activate orp_trinity && unset -f which && cd /scratch && sed -r 's|MAKEDIR :=.+|MAKEDIR := /home/orp/Oyster_River_Protocol|; s|MAKEDIR}version|MAKEDIR}/version|' /home/orp/Oyster_River_Protocol/strandeval.mk > ./strandeval.mk && chmod +x ./strandeval.mk &&  ./strandeval.mk main ASSEMBLY={} CPU=\$SLURM_CPUS_PER_TASK READ1=./input_files/unfixrm_SRO_control_R1.cor.fastq.gz READ2=./input_files/unfixrm_SRO_control_R2.cor.fastq.gz RUNOUT={/.}_orp_strandness\" " >  $JOBNAME.cmds



# submit the job array to the cluster
sbatch -a 1-$(cat $JOBNAME.cmds | wc -l)%10 --job-name=$JOBNAME --cpus-per-task=$NCORES --mem=${MEM}G --time=$WALLTIME --export=ALL,PATH=$PATH,CMDS_FILE=$JOBNAME.cmds,CONDA_NAME=base ~/bin/array.slurm


```

**Looks like we didn't assemble the reads properly!!**
Need to redo that in a strand-specific (RF) manner


```{bash funannotate-setup}
# either use conda environment or a container
CONDA_NAME="funannotate"
mamba create -n funannotate funannotate distro eggnog-mapper 


# download the interproscan container
apptainer pull docker://index.docker.io/nextgenusfs/funannotate:latest --dir $NXF_APPTAINER_CACHEDIR

# download GeneMark ES/ET (need to request a licence at http://topaz.gatech.edu/GeneMark/license_download.cg
cd ~/adna/etc/tools/Annotation
aria2c -x5 http://topaz.gatech.edu/GeneMark/tmp/GMtool_RlyKF/gmes_linux_64_4.tar.gz && tar xzfv gmes_linux_64_4.tar.gz && cd gmes_linux_64_4
wget http://topaz.gatech.edu/GeneMark/tmp/GMtool_RlyKF/gm_key_64.gz && gunzip gm_key_64.gz
ln -s  ~/etc/tools/Annotation/gmes_linux_64_4/gm_key_64 ~/etc/tools/Annotation/gmes_linux_64_4/.gm_key
# fix all perl scripts
find /home/ibar/etc/tools/Annotation/gmes_linux_64_4 -name "*.pl" | parallel "sed -ri.bak 's|/usr/bin/perl|/usr/bin/env perl|' {}"

# start and interactive job to test funannotate
start_interactive_job
CONDA_NAME="funannotate"
conda activate $CONDA_NAME
# export environment variable for gmes_petap.pl location
export GENEMARK_PATH=$(find $HOME/etc/tools/Annotation -name "gmes_petap.pl" | xargs dirname)
# add signalp to the bin or PATH (must be installed manually)
ln -s ~/scratch/tools/signalp-4.1/signalp ~/bin/
# download eggNOG data
export EGGNOG_DATA_DIR=$HOME/aDNA/tools/eggnog_dbs && mkdir -p $EGGNOG_DATA_DIR
download_eggnog_data.py
#set ENV variable for $FUNANNOTATE_DB
export FUNANNOTATE_DB=$HOME/aDNA/tools/funannotate_db
mkdir -p $FUNANNOTATE_DB
# make sure this environment variable is setup every time the environment is activated
echo "export FUNANNOTATE_DB=$HOME/aDNA/tools/funannotate_db" > $CONDA_PREFIX/etc/conda/activate.d/funannotate.sh
echo "export GENEMARK_PATH=$(find $HOME/etc/tools/Annotation -name "gmes_petap.pl" | xargs dirname)" >> $CONDA_PREFIX/etc/conda/activate.d/funannotate.sh
echo "export PATH=$PATH:\$GENEMARK_PATH" >> $CONDA_PREFIX/etc/conda/activate.d/funannotate.sh
echo "export EGGNOG_DATA_DIR=$HOME/aDNA/tools/eggnog_dbs" >> $CONDA_PREFIX/etc/conda/activate.d/funannotate.sh
echo "unset FUNANNOTATE_DB" > $CONDA_PREFIX/etc/conda/deactivate.d/funannotate.sh
echo "unset GENEMARK_PATH" >> $CONDA_PREFIX/etc/conda/deactivate.d/funannotate.sh
# reactivate environment
conda deactivate && conda activate $CONDA_NAME
#check that all modules are installed

funannotate check --show-versions

#download/setup databases to a writable/readable location
funannotate setup -d $FUNANNOTATE_DB

#run tests -- requires internet connection to download data
funannotate test -t all --cpus $SLURM_CPUS_PER_TASK

```

Now we can finally run the pipeline...

```{bash funannotate}
CONDA_NAME="funannotate"
GENOME="$HOME/adna/Oyster_QX_Disease/Nanopore_sequencing/SRO_781/Galaxy_analysis/Galaxy_Flye_assembly.fasta"
GENOME_BASENAME=$(basename -s .fasta $GENOME)
RNASEQ_DIR="$HOME/adna/Oyster_QX_Disease/QX_RNA-Seq"
export FUNANNOTATE_DB=$HOME/aDNA/tools/funannotate_db
# download Mpar proteins
aria2c -x5 -d $HOME/adna/Oyster_QX_Disease/Comparative_Genomics https://ftp.ncbi.nlm.nih.gov/genomes/all/GCA/039/583/845/GCA_039583845.1_UU_MPar_1/GCA_039583845.1_UU_MPar_1_protein.faa.gz && gunzip $HOME/adna/Oyster_QX_Disease/Comparative_Genomics/ GCA_039583845.1_UU_MPar_1_protein.faa.gz
# setup working directory
WORKDIR=/scratch/project/adna/Oyster_QX_Disease/QX_genome_annotation
mkdir -p $WORKDIR && cd $WORKDIR

NCPUS=16
MEM=64
WALLTIME="10:00:00"
JOB_NAME=funannotate-prep
echo '#!/bin/bash --login
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log'"
#SBATCH --job-name=$JOB_NAME
#SBATCH --cpus-per-task=$NCPUS
#SBATCH --mem=${MEM}G
#SBATCH --time=$WALLTIME
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
conda activate $CONDA_NAME
set -Eeo pipefail
cd \$SLURM_SUBMIT_DIR

# clean headers (not needed in our case)
# funannotate clean -i $GENOME --minlen 1000 -o $GENOME_BASENAME.cleaned.fa
# sort contigs (not needed in our case)
funannotate sort -i $GENOME --minlen 1000 -b contig -o $GENOME_BASENAME.cleaned.sorted.fa
# mask repetitive regions
funannotate mask -i $GENOME_BASENAME.cleaned.sorted.fa --cpus \$SLURM_CPUS_PER_TASK -o $GENOME_BASENAME.cleaned.sorted.masked.fa" > $JOB_NAME.slurm
# submit to the cluster
sbatch $JOB_NAME.slurm

NCPUS=16
MEM=250
WALLTIME="100:00:00"
JOB_NAME=funannotate-train
echo '#!/bin/bash --login
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log'"
#SBATCH --job-name=$JOB_NAME
#SBATCH --cpus-per-task=$NCPUS
#SBATCH --mem=${MEM}G
#SBATCH --time=$WALLTIME
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
conda activate $CONDA_NAME
set -Eeo pipefail
cd \$SLURM_SUBMIT_DIR
# train gene models based on RNA-Seq data
funannotate train -i $GENOME_BASENAME.cleaned.sorted.masked.fa -o fun --left $RNASEQ_DIR/data/unaligned_reads/NN01-567_unaligned_reads_R1.fastq.gz $RNASEQ_DIR/data/unaligned_reads/NN01-579_unaligned_reads_R1.fastq.gz $RNASEQ_DIR/data/unaligned_reads/NN01-580_unaligned_reads_R1.fastq.gz --right $RNASEQ_DIR/data/unaligned_reads/NN01-567_unaligned_reads_R2.fastq.gz $RNASEQ_DIR/data/unaligned_reads/NN01-579_unaligned_reads_R2.fastq.gz $RNASEQ_DIR/Ndata/unaligned_reads/N01-580_unaligned_reads_R2.fastq.gz --stranded RF --species 'Marteilia sydneyi' --strain SRO_781 --cpus \$SLURM_CPUS_PER_TASK --max_intronlen 10000 --trinity $RNASEQ_DIR/NN01-QX_R.Trinity.trimmed_headers.fa"  > $JOB_NAME.slurm
# submit to the cluster
sbatch $JOB_NAME.slurm

NCPUS=16
MEM=128
WALLTIME="50:00:00"
JOB_NAME=funannotate-predict
echo '#!/bin/bash --login
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log'"
#SBATCH --job-name=$JOB_NAME
#SBATCH --cpus-per-task=$NCPUS
#SBATCH --mem=${MEM}G
#SBATCH --time=$WALLTIME
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
conda activate $CONDA_NAME
set -Eeo pipefail
cd \$SLURM_SUBMIT_DIR
# predict gene models (optionally provide predicted proteins from M. para)
funannotate predict -i $GENOME_BASENAME.cleaned.sorted.masked.fa -o fun --tmpdir /scratch/project/adna/tmp --species 'Marteilia sydneyi' --strain SRO_781 --cpus \$SLURM_CPUS_PER_TASK --organism other --ploidy 2 --max_intronlen 10000 --transcript_evidence $HOME/adna/Oyster_QX_Disease/QX_RNA-Seq/NN01-QX_R.Trinity.trimmed_headers.fa --protein_evidence $HOME/adna/Oyster_QX_Disease/Comparative_Genomics/GCA_039583845.1_UU_MPar_1_protein.faa $FUNANNOTATE_DB/uniprot_sprot.fasta

# update gene predictions
funannotate update -i fun --cpus \$SLURM_CPUS_PER_TASK" >  $JOB_NAME.slurm
# submit to the cluster
sbatch $JOB_NAME.slurm

# functional annotation
--busco_db protists


```



#### stLFR Assembly

DNA sample from QX-infected oyster (NN_781) was sent to the State Agricultural Biotechnology Centre ([SABC](https://www.murdoch.edu.au/research/sabc)) at Murdoch University (led by Prof. Rajeev Varshney) for single tube long fragment read (stLFR) sequencing on an MGI DNBSEQ-T7 [@wangEfficientUniqueCobarcoding2019a]. See this paper for a comparison of sequencing and assembly strategies, including Nanopore and stLFR [@zhangComparisonDifferentSequencing2021]. 
The analysis was performed on QCIF Bunya HPC (see the [documentation](https://github.com/UQ-RCC/hpc-docs)). The raw stLFR reads need to be deconvulated based on the barcodes found in the reverse (R2) reads (a *whitelist* containing these barcodes can be found in the [stLFR_barcode_split repository](https://github.com/BGI-Qingdao/stLFR_barcode_split/blob/8a22b7f152b0f18a2d3db59733368cc4352cbc58/barcode_list.txt) or downloaded from [LRTK databases](https://drive.google.com/drive/folders/1XPW2avL_LZAt5yIh9tb35jZ5GfCSj7eQ)). Several pipelines have been developed to process such reads, as detailed below.


##### Assembly with Supernova
The [stLFR2Supernova](https://github.com/BGI-Qingdao/stlfr2supernova_pipeline) (or [stLFRdenovo](https://github.com/BGI-biotools/stLFRdenovo)) pipeline processes the raw reads, converts them to the 10X Genomics format and then uses the Supernova [@visendiNovoAssemblyLinked2022] to assemble the genome. 

```{bash stlfr-supernova}
# install stLFRdenovo pipeline and dependencies
cd ~/etc/tools/
wget https://github.com/BGI-biotools/stLFRdenovo/releases/download/v1.0.5/stLFRdenovo-v1.0.5.tar.gz
tar xzf stLFRdenovo-v1.0.5.tar.gz
WHITELIST=$(find ~/etc/tools -name "barcode_list.txt")



```

##### Assembly with STLFRassembly
Another approach that may be more suitable for our scenario is implemented in pipelines that attempt to assemble metagenomes from the stLFR reads. 

* The [MetaSTHD Assembly Pipeline](https://github.com/sumoii/STLFRassembly) uses several tools, such as [MetaWRAP](https://github.com/bxlab/metaWRAP) [@uritskiyMetaWRAPFlexiblePipeline2018], [WENGAN](https://github.com/adigenova/wengan) [@digenovaEfficientHybridNovo2021] and [cloudSPAdes](https://github.com/ablab/spades/tree/cloudspades-paper)[@tolstoganovCloudSPAdesAssemblySynthetic2019] to perform this task.
* [LRTK](https://github.com/ericcombiolab/LRTK) is a similar pipeline that seems much more complete and mature [@yangLRTKPlatformAgnostic2022].
* [Athena_meta](https://github.com/abishara/athena_meta) is a linked-reads metagenomes assembly tool. It can be installed via `conda` or as a Docker container `abishara/athena-meta-flye-docker` [@bisharaHighqualityGenomeSequences2018].  
* [MetaTrass](https://github.com/BGI-Qingdao/MetaTrass) is a linked-reads metagenomes assembly tool for the human gut microbiome [@qiMetaTrassHighqualityMetagenome2022] 

```{bash stlr-LRTK}
CONDA_NAME=lrtk
# install LRTK Assembly Pipeline and dependencies
mamba create -n $CONDA_NAME chopper lrtk pigz
# Athena_meta can be installed and run from a Docker container (abishara/athena-meta-flye-docker)
# convert to a unified linked-read format
WORKDIR=/scratch/project/adna/Oyster_QX_Disease/stLFR_Murdoch
mkdir -p $WORKDIR/SRO-lrtk/converted_reads && cd $WORKDIR/SRO-lrtk
cd $WORKDIR/SRO-lrtk
# download and extract barcode whitelist
wget -qO- https://raw.githubusercontent.com/BGI-Qingdao/stLFR_barcode_split/8a22b7f152b0f18a2d3db59733368cc4352cbc58/barcode_list.txt | gawk '{printf ">%s\n%s\n", $2, $1}' > $WORKDIR/stLFR_barcodes.fa
# find ~/etc/tools -iwholename "*stLFR*/barcode_list.txt"
WHITELIST=$WORKDIR/stLFR_barcodes.fa
# WHITELIST=$(find $WORKDIR/SRO-lrtk -name "white_list_stlfr_barcode.fa" |  head -n 1)
# download host reference
rclone copyto -P GURS_shared:Oyster_QX_Disease/Sro_reference_concatenated.fasta /scratch/project/adna/Oyster_QX_Disease//Sro_reference_concatenated.fasta
HOST=/scratch/project/adna/Oyster_QX_Disease/Sro_reference_concatenated.fasta

NCPUS=16
MEM=64
WALLTIME="20:00:00"
JOB_NAME=SRO-LRTK-convert
echo '#!/bin/bash --login
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log'"
#SBATCH --job-name=$JOB_NAME
#SBATCH --cpus-per-task=$NCPUS
#SBATCH --mem=${MEM}G
#SBATCH --time=$WALLTIME
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
conda activate $CONDA_NAME
set -Eeo pipefail
cd \$SLURM_SUBMIT_DIR
pigz -cd $WORKDIR/Oyster_781.2.R1.fq.gz > Oyster_781.2.R1.fq
pigz -cd $WORKDIR/Oyster_781.2.R2.fq.gz > Oyster_781.2.R2.fq
bwa index $WHITELIST
bwa index $HOST
srun lrtk FQCONVER -I1 Oyster_781.2.R1.fq -I2 Oyster_781.2.R2.fq -IT stLFR \
-O1 $WORKDIR/SRO-lrtk/converted_reads/Oyster_781.2.R1.non-host.fq \
-O2 $WORKDIR/SRO-lrtk/converted_reads/Oyster_781.2.R2.non-host.fq \
-BW $WHITELIST -T \$SLURM_CPUS_PER_TASK -HD $HOST \
-F Yes -S Yes -G metagenome
rm $WORKDIR/Oyster_781.2.R*.fq" > $JOB_NAME.slurm
# submit the job 
sbatch $JOB_NAME.slurm

# Barcode-aware alignment to the host reference

lrtk ALIGN -FQ1 /path_to/IN_FQ1 -FQ2 /path_to/IN_FQ2 -R /path_to/REFERENCE -O /path_to/OUT_BAM -RG "@RG\tID:Example\tSM:Example" -P 10x -T 4



```


```{bash stlr-metaSTHD}
# install MetaSTHD Assembly Pipeline and dependencies
# install wengan
cd ~/etc/tools/
wget https://github.com/adigenova/wengan/releases/download/v0.2/wengan-v0.2-bin-Linux.tar.gz
tar xzf wengan-v0.2-bin-Linux.tar.gz
# install STLFRassembly
git clone https://github.com/sumoii/STLFRassembly.git
# fix shebang
ls -1 ~/etc/tools/STLFRassembly/shellfile/bin/*.sh | parallel "sed -ri.bak '1s|#.+|#!/usr/bin/bash|; s/mkdir/mkdir -p/g'; s/ln -s/ln -sf/g"
sed '138i scriptpath=`echo $(readlink -f ${BASH_SOURCE[0]})`' | sed '139i binpath=${scriptpath%/*}' | sed -r 's|sh Step|sh $binpath/Step|g; s|source Step|source $binpath/Step|g; s/allmethod=right/allmethod=$OPTARG/' ~/etc/tools/STLFRassembly/shellfile/bin/run.sh
# make running script executable
chmod +x /home/ibar/etc/tools/STLFRassembly/shellfile/bin/run.sh

CONDA_NAME=stlfr-meta
mamba create -n $CONDA_NAME perl blas=2.5=mkl ursky::metawrap-mg=1.3.2 quast wtdbg 
conda activate $CONDA_NAME
# install cloudspades
aria2c -x5 https://github.com/ablab/spades/archive/refs/tags/cloudspades-paper.tar.gz
tar xzf cloudspades-paper.tar.gz
cd spades-cloudspades-paper/assembler
./spades_compile.sh

# conda activate $CONDA_NAME

# download taxonomy databases
# TX_DB_DIR="/project/aerc/ncbi/new_taxdump"
# mkdir -p $TX_DB_DIR
# cd $TX_DB_DIR
# curl -L https://ftp.ncbi.nih.gov/pub/taxonomy/new_taxdump/new_taxdump.tar.gz | tar xzf -

# setup working directory
WORK_DIR=/scratch/project/adna/Oyster_QX_Disease/stLFR_Murdoch
cd $WORK_DIR
rclone copy -P --exclude "*unmapped_reads.fastq.gz" GURS_shared:Oyster_QX_Disease/Nanopore_sequencing/fastq_files ./Nanopore_reads
# filter only long reads (>5kb)
echo '#!/bin/bash --login
#SBATCH --job-name=ont-filter
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log
#SBATCH --cpus-per-task=12
#SBATCH --mem=48G
#SBATCH --time=10:00:00
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
set -Eeo pipefail
cd $SLURM_SUBMIT_DIR
srun zcat Nanopore_reads/*.fastq.gz | chopper -q 10 -l 5000 --threads 12 | gzip > SRO_781_ONT.filtered.fastq.gz' > SRO_filter_ont.slurm
sbatch SRO_filter_ont.slurm

WHITELIST=$(find ~/etc/tools -name "barcode_list.txt")
NCPUS=36
MEM=500
MODEL=M # or A
echo '#!/bin/bash --login
#SBATCH --job-name=SRO-stlfr-meta
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --output=%x.%j.log'"
#SBATCH --cpus-per-task=$NCPUS
#SBATCH --mem=${MEM}G
#SBATCH --time=200:00:00
#SBATCH --account=a_agri_genomics
#SBATCH --partition=general

source ~/.bashrc
conda activate $CONDA_NAME
set -Eeo pipefail
cd \$SLURM_SUBMIT_DIR
export wengan=$HOME/etc/tools/wengan-v0.2-bin-Linux/wengan.pl
srun bash $HOME/etc/tools/STLFRassembly/shellfile/bin/run.sh -l \
$PWD/SRO_781_ONT.filtered.fastq.gz -1 $PWD/Oyster_781.2.R1.fq.gz \
-2 $PWD/Oyster_781.2.R2.fq.gz -t $NCPUS -c /home/ibar/etc/tools/spades-cloudspades-paper/assembler/spades.py -L -w wtdbg2 \
-g $MEM -m $MODEL -M Binning -b metaWRAP -A false -w $WHITELIST" > SRO_stlfr_meta.slurm
# submit the job 
sbatch SRO_stlfr_meta.slurm


```

## General information

This document was last updated at `r Sys.time()` using R Markdown (built with `r R.version.string`). Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. It is especially powerful at authoring documents and reports which include code and can execute code and use the results in the output. For more details on using R Markdown see <http://rmarkdown.rstudio.com> and [Rmarkdown cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf).

------------------------------------------------------------------------

## Bibliography